---
title: "Further applications of Paralog Annotation"
# output: rmarkdown::github_document
output: 

  # word_document:
  #  reference_docx: template.docx

  pdf_document:
    toc: true
    number_sections: true
    fig_caption: yes
    
editor_options: 
  chunk_output_type: inline
bibliography: bibliography.bib
always_allow_html: yes
---

<!--Load Packages and function-->
<!-- ```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide'} -->
```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
source("src/Paralogous_var_align_functions.R")
```
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```
\newpage
# Introduction
With the advancements of sequencing technology, new potential variants are being discovered constantly. However to be able to identify said variants as being pathogenic requires supporting evidence, which does not always exists especially if the variant is novel. Several commonly used bioinformatic tools for predicting the functional effect of variants, such as SIFT and REVEL [@sim2012sift,@ioannidis2016revel], utilize information from conservation and/or changes in biophysiocochemical properties of the substituting amino acid. However these predictions cannot be considered as sufficient evidence for pathogenicity if predicted as such and need to be used with care, especially in a clinical context, as results typically have low specificity [@flanagan2010using].

Previously Ware *et al.* developed the concept of **Paralogue Annotation** [@Ware2012; @Walsh2014], which utilizes known variant information from paralogues (evolutionarily related genes arising from gene duplication from the same species) to help classify pathogenic variants. They verified its use in LQT syndrome (LQTS) and Brugada syndrome associated genes on variants acquired from patient cohorts.

Here **Paralogue Annotation** is tested further using a Pathogenic/Likely Pathogenic (P/LP) and Benign/Likely Benign (B/LB) varaint dataset acquired from Clinvar. These variants covered a wider range of genes, and were used to demonstrate and validate the application of **Paralogue Annotation** in comparison to common variant interpretation tools used today.

\newpage
# Material and Methods

## Datasets
Clinvar Likely Pathogenic/Pathogenic and Likely Benign/Benign variant vcf files were extracted and downloaded via the method developed by Zhang *et al.* [@Zhang2017]. Variants that were positioned in any of the following 8 sarcomeric: MYH7; MYBPC3; TNNT2; TPM1; MYL2; MYL3; TNNI3; and ACTC1, were also subsetted for further analysis as the case example and used for etiological fraction calculations.

Clinical significance definitions of variants - Pathogenic; Likely Pathogenic; Likely Benign; Benign - were defined according to Clinvar [@landrum2015clinvar]. Variants of unknown significance were not used in the study.

For the case and control study, HCM diseased cohorts data were gathered from the OMGL and LMM datasets, which were taken from previous @Walsh2017 publication. Data from ExAC were used as the control cohort @Lek2016.

## Annotation of variants and transfer of annotations across paralogues

The data was prepared by removing frameshift mutations

## Statistical measures
For a pathogenic paralogue alignment:

true positive (TP) = pathogenic query variant with a paralogous pathogenic hit;

false positive (FP) = benign query variant with a paralogous pathogenic hit;

false negative (FN) = pathogenic query variant with no paralogous pathogenic hit;

and true negative (TN) = benign query variant with no paralogous pathogenic hit.\newline


Likewise for a benign paralogous alignment:

TP = benign query variant with a paralogous benign hit; 

FP = pathogenic query variant with a paralogous benign hit; 

FN = benign query variant with no paralogous benign hit; 

and TN = pathogenic query variant with no paralogous benign hit.\newline

Positive Predictive Values (PPV) and Sensitivties are therefore calculated by:

$$PPV = \frac{TP}{TP+FP}$$
$$Sensitivity = \frac{TP}{TP+FN}$$

P values are calculated via a Fisher's exact test on a 2x2 contingency table tabulating the number of pathogenic and benign variants of interest and how many of those are predicted as pathogenic or benign.

## Calculation of Etiological Fractions

Odds ratios (OR) are calculated by:

$$ OR =  \frac{(a/b)}{(c/d)} $$

where:

$a =$ number of variants predicted to be pathogenic by paralogue annotation in the diseased cohort

$b =$ number of variants not predicted to be pathogenic by paralogue annotation in the diseased cohort

$c =$ number of varaints predicted to be pathogenic by paralogue annotation in the control cohort

$d =$ number of variants not predicted to be pathogenic by paralogue annotation in the control cohort

and the 95% confidence intervals for OR values are calculated according to @altman1991practical via:

$$ 95\% \ CI = [e^{\ln(OR) - 1.96 \cdot SE(\ln(OR))},e^{\ln(OR) + 1.96 \cdot SE(\ln(OR))}] $$
where:

$$ SE(\ln(OR)) =  \sqrt{\frac{1}{a}+\frac{1}{b}+\frac{1}{c}+\frac{1}{d}} $$

Etiological Fractions (EF) can then be calculated by:

$$ EF = \frac{OR - 1}{OR} $$
and the 95% confidence intervals for EF values are calculated according to @hildebrandt2006calculating via:

$$ 95\% \ CI = [\widehat{EF} - z_{1-\frac{a}{2}} \sqrt{\widehat{VAR}(\widehat{\Phi})}, \ min( \widehat{EF} + z_{1-\frac{a}{2}} \sqrt{\widehat{VAR}(\widehat{\Phi})}, \ 1)]  $$
where the asymptomatic variance estimator of $\widehat{\Phi}$ is given by:

$$ \widehat{VAR}(\widehat{\Phi}) = \widehat{\Phi}^2 [\frac{1-\widehat{\pi}_{0}}{N_0\widehat{\pi}_{0}}+\frac{1-\widehat{\pi}_{1}}{N_1\widehat{\pi}_{1}}]$$
and:
$\widehat{\pi}_{1} = a/(b + a)$
$\widehat{\pi}_{0} = c/(d + c)$ 
$N_0 = c + d$ 
$N_1 = a + b$ 
$\widehat{\Phi}^2 = (\widehat{\pi}_{0}/\widehat{\pi}_{1})^2$

NOTE ACCORDING TO https://www.medcalc.org/calc/odds_ratio.php - "Where zeros cause problems with computation of the odds ratio or its standard error, 0.5 is added to all cells (a, b, c, d) (Pagano & Gauvreau, 2000; Deeks & Higgins, 2010)."

## Paralogue stats
The additional statistics were calculated by programmatically extracting the genes of interest (using `src/check_what_clinvar_genes.py` and `src/Find_unique_genes.py`) and then retrieving relevant information manually from [Ensembl's Bioimart](https://www.ensembl.org/biomart)

Alternatively, this can be reproduced using biomaRt package
                 
## Para-Z scores
For the para-z scores, will need to extract amino acid position from VEP output as well. Then look up the gene in question in para-z score folder, and using the position identify the para-z score. From my understanding, the para-z score is the same across aligned amino acids in the same gene family. Therefore, we could use a cut-off threshold to further improve our confidence in calling variants pathogenic etc. We could also then calculate ROC curves by altering the cut-off to see how that affects sensitivity/PPV.

All available para-z scores were retreived from https://git-r3lab.uni.lu/genomeanalysis/paralogs/tree/master/data [@Lal2017]. Para-Z score cutoff thresholds were used to remove any annotation alignments in question. Amino Acid positions that had a para-z score below the chosen cutoff threshold were not used for annotations.

\newpage

## Comparison to SIFT/REVEL scores
SIFT and Mcap scores were retreived using ANNOVAR with dbnsfp30a [@wang2010annovar,@liu2016dbnsfp]. Revel scores were downloaded from https://sites.google.com/site/revelgenomics/downloads [@ioannidis2016revel].

Revel did not provide scores for all available variants in the dataset and therefore to ensure the same variants were being annotated by all tools

# Results and Discussion

## Paralogue stats

```{r paralog_dist, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7,fig.height=6,fig.cap="\\label{fig:paralog_dist}Distribution of genes with paralogues by the number of paralogues they're related to"}
ensembl = useMart("ENSEMBL_MART_ENSEMBL", host = "jul2018.archive.ensembl.org")
ens_human <- useDataset("hsapiens_gene_ensembl",mart=ensembl)
protein_genes_w_paralogues = getBM(attributes = c("external_gene_name", "hsapiens_paralog_associated_gene_name"), filters = c("biotype", "with_hsapiens_paralog"), values = list("protein_coding", TRUE), mart = ens_human)
protein_genes_w_paralogues_genes = as.vector(unique(protein_genes_w_paralogues$external_gene_name))

# eg = bitr(protein_genes_w_paralogues_genes, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
# ego = groupGO(gene = eg$ENTREZID, OrgDb = org.Hs.eg.db, keyType = "ENTREZID", readable = TRUE, ont = "MF", level = 2)
# barplot(ego)


# length(protein_genes_w_paralogues)
protein_genes_w_paralogues_wide = plyr::count(protein_genes_w_paralogues, "external_gene_name")
protein_genes_w_paralogues_wide$paralogue = "Yes"
colnames(protein_genes_w_paralogues_wide)[2] = "number_of_paralogues"
par(mfrow=c(2,2))
# hist(protein_genes_w_paralogues_wide$number_of_paralogues, breaks = seq(0,50,1), xlab = "number of paralogues", ylab = "number of genes", main="A) Overall distribution of genes with paralogues", cex.main = 0.7)

protein_genes_wo_paralogues = getBM(attributes = c("external_gene_name", "hsapiens_paralog_associated_gene_name"), filters = c("biotype", "with_hsapiens_paralog"), values = list("protein_coding", FALSE), mart = ens_human)
protein_genes_wo_paralogues = protein_genes_wo_paralogues[!(protein_genes_wo_paralogues$external_gene_name %in% protein_genes_w_paralogues_genes),]
protein_genes_wo_paralogues_genes = as.vector(unique(protein_genes_wo_paralogues$external_gene_name))

# length(protein_genes_wo_paralogues)
protein_genes_wo_paralogues_wide = protein_genes_wo_paralogues
protein_genes_wo_paralogues_wide[is.na(protein_genes_wo_paralogues_wide)] = 0
protein_genes_wo_paralogues_wide$paralogue = "No"
colnames(protein_genes_wo_paralogues_wide)[2] = "number_of_paralogues"
protein_genes_wo_paralogues_wide = protein_genes_wo_paralogues_wide[!(protein_genes_wo_paralogues_wide$external_gene_name %in% protein_genes_w_paralogues_wide$external_gene_name),]


#testing to see how many genes in the clinvar dataset have paralogues
clinvar_P_LP_tableized = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
clinvar_P_LP_tableized_genes = as.vector(unique(clinvar_P_LP_tableized$SYMBOL))
clinvar_P_LP_vcf_tableized = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
clinvar_P_LP_vcf_tableized_genes = as.vector(unique(clinvar_P_LP_vcf_tableized$GENEINFO))
clinvar_P_LP_vcf_tableized_genes = unique(unlist(sapply(clinvar_P_LP_vcf_tableized_genes, strsplit, "\\|")))
clinvar_P_LP_vcf_tableized_genes = sapply(strsplit(clinvar_P_LP_vcf_tableized_genes, "\\:"), "[", 1)

clinvar_B_LB_tableized = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
clinvar_B_LB_tableized_genes = as.vector(unique(clinvar_B_LB_tableized$SYMBOL))
clinvar_B_LB_vcf_tableized = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
clinvar_B_LB_vcf_tableized_genes = as.vector(unique(clinvar_B_LB_vcf_tableized$GENEINFO))
clinvar_B_LB_vcf_tableized_genes = unique(unlist(sapply(clinvar_B_LB_vcf_tableized_genes, strsplit, "\\|")))
clinvar_B_LB_vcf_tableized_genes = sapply(strsplit(clinvar_B_LB_vcf_tableized_genes, "\\:"), "[", 1)

#total clinvar genes in both P/LP and B/LB
all_clinvar_genes = unique(c(clinvar_P_LP_tableized_genes, clinvar_B_LB_tableized_genes))

#ALL clinvar genes including VUS, not just P/LP and B/LB
full_clinvar_tableized = read.csv(file="data/clinvar/clinvar_20190114.vcf_tableized_org", sep = "\t", header = TRUE, stringsAsFactors = FALSE)

overlapped_genes = vector()
non_overlapped_genes = vector()
for (x in unique(full_clinvar_tableized$GENEINFO)){
  if (grepl("\\|", x)){
    overlapped_genes = c(overlapped_genes, x)
  } else {
    non_overlapped_genes = c(non_overlapped_genes, x)
  }
}
overlapped_genes = unique(unlist(sapply(overlapped_genes, strsplit, "\\|")))
overlapped_genes = sapply(strsplit(overlapped_genes, "\\:"), "[", 1)
non_overlapped_genes = unique(unlist(sapply(non_overlapped_genes, strsplit, "\\|")))
non_overlapped_genes = sapply(strsplit(non_overlapped_genes, "\\:"), "[", 1)
full_genes = unique(c(overlapped_genes, non_overlapped_genes)) 
full_genes_w_paralogues = intersect(full_genes, protein_genes_w_paralogues_genes)


# Total protein genes
all_protein_genes = dplyr::bind_rows(protein_genes_wo_paralogues_wide, protein_genes_w_paralogues_wide)
his = ggplot(all_protein_genes, aes(x=number_of_paralogues, fill=paralogue)) + geom_histogram(binwidth = 1, colour = "black") + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    plot.title = element_text(hjust = 0.5)) +labs(title = "Overall distribution of human protein coding genes in Ensembl and their associated paralogues", 
    x = "number of paralogues", y = "number of genes", 
    fill = "Has paralogues")+labs(title = "Overall distribution of human protein coding
genes in Ensembl and their associated paralogues")+ylim(0,6000) + theme(axis.title = element_text(size = 15), 
    axis.text = element_text(size = 15), 
    plot.title = element_text(size = 15), 
    legend.text = element_text(size = 12), 
    legend.title = element_text(size = 12))
ggsave("images/Total_genes_paralogs.png")
# Subset disease genes only
disease_genes = vector()
for (gene in clinvar_P_LP_tableized_genes){
  # print(gene)
  if (gene %in% all_protein_genes$external_gene_name){
    disease_genes = c(gene, disease_genes)
  }
}
# Disease genes
disease_genes_wide = all_protein_genes[all_protein_genes$external_gene_name %in% disease_genes,]
his2 = ggplot(disease_genes_wide, aes(x=number_of_paralogues, fill=paralogue)) + geom_histogram(binwidth = 1, colour = "black") + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    plot.title = element_text(hjust = 0.5)) +labs(title = "Distribution of human protein coding genes involved in disease in Ensembl and their associated paralogues", 
    x = "number of paralogues", y = "number of genes", 
    fill = "Has paralogues")+labs(title = "Distribution of human protein coding genes involved
in disease in Ensembl and their associated paralogues") + theme(axis.title = element_text(size = 15), 
    axis.text = element_text(size = 15), 
    plot.title = element_text(size = 15), 
    legend.text = element_text(size = 12), 
    legend.title = element_text(size = 12))
ggsave("images/Disease_genes_paralogs.png")

#combined total protein and disease gene dsitributions

disease_genes_wide$disease = "Yes"
all_protein_genes$disease = "No"

total_plot_data = rbind(all_protein_genes, disease_genes_wide)
total_plot_data[total_plot_data$disease == "No" & total_plot_data$paralogue == "No", "colour"] = "no_dis_no_para"
total_plot_data[total_plot_data$disease == "No" & total_plot_data$paralogue == "Yes", "colour"] = "no_dis_yes_para"
total_plot_data[total_plot_data$disease == "Yes" & total_plot_data$paralogue == "No", "colour"] = "yes_dis_no_para"
total_plot_data[total_plot_data$disease == "Yes" & total_plot_data$paralogue == "Yes", "colour"] = "yes_dis_yes_para"


cols = c("no_dis_no_para" = "orange", "yes_dis_no_para" = "red", "no_dis_yes_para" = "purple", "yes_dis_yes_para" = "blue")
dens = ggplot(total_plot_data, aes(x=number_of_paralogues)) + 
  geom_histogram(data = total_plot_data[total_plot_data$disease=="No" & total_plot_data$paralogue == "No",], aes(fill = "no_dis_no_para"), alpha = 0.4, binwidth = 1) +
  geom_histogram(data = total_plot_data[total_plot_data$disease=="Yes" & total_plot_data$paralogue == "No",], aes(fill = "yes_dis_no_para"), alpha = 0.9, binwidth = 1) +
  geom_histogram(data = total_plot_data[total_plot_data$disease=="No" & total_plot_data$paralogue == "Yes",], aes(fill = "no_dis_yes_para"), alpha = 0.4, binwidth = 1) +
  geom_histogram(data = total_plot_data[total_plot_data$disease=="Yes" & total_plot_data$paralogue == "Yes",], aes(fill = "yes_dis_yes_para"), alpha = 0.9, binwidth = 1) +
  scale_fill_manual(values = cols, labels = c("All genes, no paralogues", "All genes, has paralogues", "Clinvar disease genes, no paralogues",  "Clinvar disease genes, has paralogues")) +
  theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    plot.title = element_text(hjust = 0.5)) +labs(title = "Distribution of human protein coding genes involved in disease in Ensembl and their associated paralogues", 
    x = "number of paralogues", y = "number of genes", 
    fill = "")+labs(title = "Distribution of human protein coding genes involved
in disease in Ensembl and their associated paralogues") + theme(axis.title = element_text(size = 15), 
    axis.text = element_text(size = 15), 
    plot.title = element_text(size = 15), 
    legend.text = element_text(size = 12), 
    legend.title = element_text(size = 12))+labs(fill = NULL)
ggsave("images/Total_and_Disease_genes_paralogs.png", height = 5, width = 10)

# dens2 = ggplot(disease_genes_wide, aes(x=number_of_paralogues)) + geom_density()

#see if Patho and Benign sets have same distribution of variants in genes with paralogues
clinvar_P_LP_genes_w_paralogs = vector()
clinvar_P_LP_genes_wo_paralogs = vector()
clinvar_B_LB_genes_w_paralogs = vector()
clinvar_B_LB_genes_wo_paralogs= vector()
for (gene in clinvar_P_LP_tableized_genes){
  # print(gene)
  if (gene %in% protein_genes_w_paralogues_genes){
    clinvar_P_LP_genes_w_paralogs = c(gene, clinvar_P_LP_genes_w_paralogs)
  } else {
    clinvar_P_LP_genes_wo_paralogs = c(gene, clinvar_P_LP_genes_wo_paralogs)
  }
}
clinvar_P_LP_genes_w_paralogs_wide = protein_genes_w_paralogues_wide[protein_genes_w_paralogues_wide$external_gene_name %in% clinvar_P_LP_genes_w_paralogs,]
# hist(clinvar_P_LP_genes_w_paralogs_wide$number_of_paralogues, breaks = seq(0,50,1), xlab = "number of paralogues", ylab = "number of genes", main="B) Distribution of genes with pathogenic/likely pathogenic variants", cex.main = 0.7)


for (gene in clinvar_B_LB_tableized_genes){
  # print(gene)
  if (gene %in% protein_genes_w_paralogues_genes){
    clinvar_B_LB_genes_w_paralogs = c(gene, clinvar_B_LB_genes_w_paralogs)
  } else {
    clinvar_B_LB_genes_wo_paralogs = c(gene, clinvar_B_LB_genes_wo_paralogs)
  }
}
clinvar_B_LB_genes_w_paralogs_wide = protein_genes_w_paralogues_wide[protein_genes_w_paralogues_wide$external_gene_name %in% clinvar_B_LB_genes_w_paralogs,]
# hist(clinvar_B_LB_genes_w_paralogs_wide$number_of_paralogues, breaks = seq(0,50,1), xlab = "number of paralogues", ylab = "number of genes", main="C) Distribution of genes with benign/likely benign variants", cex.main = 0.7)

clinvar_P_LP_variants_wo_paralogs = vector()
for (variant in clinvar_P_LP_tableized$SYMBOL){
  if (variant %in% protein_genes_wo_paralogues_genes){
    clinvar_P_LP_variants_wo_paralogs = c(variant, clinvar_P_LP_variants_wo_paralogs)
  }
}

clinvar_B_LB_variants_wo_paralogs = vector()
for (variant in clinvar_B_LB_tableized$SYMBOL){
  if (variant %in% protein_genes_wo_paralogues_genes){
    clinvar_B_LB_variants_wo_paralogs = c(variant, clinvar_B_LB_variants_wo_paralogs)
  }
}

ks = ks.test(clinvar_P_LP_genes_w_paralogs_wide$number_of_paralogues, clinvar_B_LB_genes_w_paralogs_wide$number_of_paralogues)

# cap = "\\label{fig:paralog_dist}Distribution of genes with paralogues by the number of paralogues they're related to"

#total number of paralogs in clinvar set
all_clinvar_gene_wo_paralogs = unique(c(clinvar_P_LP_genes_wo_paralogs, clinvar_B_LB_genes_wo_paralogs))
all_clinvar_gene_w_paralogs = unique(c(clinvar_P_LP_genes_w_paralogs, clinvar_B_LB_genes_w_paralogs))
write.table(all_clinvar_gene_w_paralogs,file="data/clinvar/all_clinvar_gene_w_paralogs.txt", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)
```

According to ensembl, `r length(protein_genes_w_paralogues_genes)` protein coding genes are defined to have paralogues. While `r length(protein_genes_wo_paralogues_genes)` protein coding genes did not have paralogues. Of those genes with paralogues (**fig.** \ref{fig:paralog_dist}a) the mean had `r round(mean(protein_genes_w_paralogues_wide$number_of_paralogues), 3)` paralogues with a standard deviation of `r round(sd(protein_genes_w_paralogues_wide$number_of_paralogues), 3)`. The maximum number of paralogues a gene had was `r max(protein_genes_w_paralogues_wide$number_of_paralogues)`. The zinc finger family of proteins have highest number of paralogues

In the clinvar pathogenic and likely pathogenic dataset, there were `r length(unique(clinvar_P_LP_tableized$ID))` variants from `r length(clinvar_P_LP_tableized_genes)` genes. `r length(clinvar_P_LP_genes_wo_paralogs)` of these did not have paralogs and therefore the `r length(clinvar_P_LP_variants_wo_paralogs)` variants lying within these genes were not used for annotation, leaving `r length(clinvar_P_LP_tableized$CHROM) - length(clinvar_P_LP_variants_wo_paralogs)` for use in the analysis. The distribution of number of paralogues for these set of genes is shown in **fig.** \ref{fig:paralog_dist}b. The mean number of paralogues was `r round(mean(clinvar_P_LP_genes_w_paralogs_wide$number_of_paralogues), 3)` with a standard deviation `r round(sd(clinvar_P_LP_genes_w_paralogs_wide$number_of_paralogues), 3)`.

For variants in the clinvar benign and likely benign dataset, there were `r length(clinvar_B_LB_tableized$CHROM)` variants from `r length(clinvar_B_LB_tableized_genes)` genes. `r length(clinvar_B_LB_tableized$CHROM) - length(clinvar_B_LB_variants_wo_paralogs)` variants resided in `r length(clinvar_B_LB_genes_w_paralogs)` genes with paralogs. Their respective distribution is shown in **fig.** \ref{fig:paralog_dist}c, with a mean of `r round(mean(clinvar_P_LP_genes_w_paralogs_wide$number_of_paralogues), 3)` paralogues and a standard deviation of `r round(sd(clinvar_P_LP_genes_w_paralogs_wide$number_of_paralogues), 3)`.

Performing a simple kolmogorov smirnov test between the distribution of pathogenic variants in genes with paralogues and benign variants shows a p-value of `r ks$p.value` suggesting that the null hypothesis of the distributions being identical cannot be rejected. From this, there appears to be no statistical difference between pathogenic variants being more likely to lie in genes that have more paralogs compared to benign, at least in regards to the definitions of clinical significance made by clinvar.

## Annotation of Clinvar
```{r clinvar_table, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE, fig.width=7,fig.height=6,fig.cap="\\label{fig:clinvar_table}Total number of known P/LP variants and known B/LB variants predicted to be pathogenic"}
# library(knitr)
#noQC
#pathogenic to pathogenic
p.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# p.paralogous_var_align = Paralogous_var_align("/media/nick/Data/Users/N/Documents/PhD/Paralogues/data_files/clinvar_20171029_onlyPathogenic.out_paraloc_paralogs2.noQC", "/media/nick/Data/Users/N/Documents/PhD/Paralogues/data_files/clinvar_20171029_onlyPathogenic.out_paraloc_tableized")
#benign to pathogenic
b.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
#pathogenic to benign
ptob.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")
#benign to benign
btob.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")
#stats
conf = conf_matrix(p.paralogous_var_align$true_num_of_paralog_anno, p.paralogous_var_align$input_variants, b.paralogous_var_align$true_num_of_paralog_anno, b.paralogous_var_align$input_variants)
conf_b = conf_matrix_benign(ptob.paralogous_var_align$true_num_of_paralog_anno, ptob.paralogous_var_align$input_variants, btob.paralogous_var_align$true_num_of_paralog_anno, btob.paralogous_var_align$input_variants)

genes_before_PA = length(unique(p.paralogous_var_align$Total_paralog_annotations$Gene))
write.table(unique(p.paralogous_var_align$Total_paralog_annotations$Gene), file = "data/clinvar/genes_before_paralogue_annotation.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)
genes_with_PA_noQC = length(unique(p.paralogous_var_align$Total_paralog_annotations$SYMBOL))
write.table(unique(p.paralogous_var_align$Total_paralog_annotations$SYMBOL), file = "data/clinvar/genes_with_paralogue_annotation_noQC.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)

#finding out what genes the variants are in
# var_in_genes = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")#,Overlap = 0)
var_in_genes_unique_var = data.frame(table(p.paralogous_var_align$Unique_variant_gene_annotations$Gene))
p.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
p.var_in_genes_unique_var = p.var_in_genes_unique_var[p.var_in_genes_unique_var$Var1 %in% clinvar_P_LP_tableized_genes,]
write.table(p.var_in_genes_unique_var, file = "data/clinvar/genes_with_number_of_annotations_noQC.csv", na="", row.names = FALSE, col.names = TRUE, sep = ",", quote=FALSE)
not_var_in_genes_unique_var = data.frame(table(p.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
p.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
p.not_var_in_genes_unique_var = p.not_var_in_genes_unique_var[p.not_var_in_genes_unique_var$Var1 %in% clinvar_P_LP_tableized_genes,]
length(intersect(p.var_in_genes_unique_var$Var1, p.not_var_in_genes_unique_var$Var1))

#top 5 genes where PA performs best with clinvar data
COL3A1_annotations = p.paralogous_var_align$Total_paralog_annotations[p.paralogous_var_align$Total_paralog_annotations$Gene == "COL3A1" & !is.na(p.paralogous_var_align$Total_paralog_annotations$ID.y),]
COL3A1_annotations_genes = data.frame(table(COL3A1_annotations$SYMBOL))
COL3A1_annotations_genes = arrange(COL3A1_annotations_genes, desc(Freq))
COL3A1_annotations_genes$Gene = "COL3A1"
COL4A5_annotations = p.paralogous_var_align$Total_paralog_annotations[p.paralogous_var_align$Total_paralog_annotations$Gene == "COL4A5" & !is.na(p.paralogous_var_align$Total_paralog_annotations$ID.y),]
COL4A5_annotations_genes = data.frame(table(COL4A5_annotations$SYMBOL))
COL4A5_annotations_genes = arrange(COL4A5_annotations_genes, desc(Freq))
COL4A5_annotations_genes$Gene = "COL4A5"
SCN1A_annotations = p.paralogous_var_align$Total_paralog_annotations[p.paralogous_var_align$Total_paralog_annotations$Gene == "SCN1A" & !is.na(p.paralogous_var_align$Total_paralog_annotations$ID.y),]
SCN1A_annotations_genes = data.frame(table(SCN1A_annotations$SYMBOL))
SCN1A_annotations_genes = arrange(SCN1A_annotations_genes, desc(Freq))
SCN1A_annotations_genes$Gene = "SCN1A"
COL1A1_annotations = p.paralogous_var_align$Total_paralog_annotations[p.paralogous_var_align$Total_paralog_annotations$Gene == "COL1A1" & !is.na(p.paralogous_var_align$Total_paralog_annotations$ID.y),]
COL1A1_annotations_genes = data.frame(table(COL1A1_annotations$SYMBOL))
COL1A1_annotations_genes = arrange(COL1A1_annotations_genes, desc(Freq))
COL1A1_annotations_genes$Gene = "COL1A1"
COL1A2_annotations = p.paralogous_var_align$Total_paralog_annotations[p.paralogous_var_align$Total_paralog_annotations$Gene == "COL1A2" & !is.na(p.paralogous_var_align$Total_paralog_annotations$ID.y),]
COL1A2_annotations_genes = data.frame(table(COL1A2_annotations$SYMBOL))
COL1A2_annotations_genes = arrange(COL1A2_annotations_genes, desc(Freq))
COL1A2_annotations_genes$Gene = "COL1A2"
top_5_paralogues = rbind(COL3A1_annotations_genes, COL4A5_annotations_genes, SCN1A_annotations_genes, COL1A1_annotations_genes, COL1A2_annotations_genes)
top_5_paralogues = top_5_paralogues[,c(3,1,2)]
names(top_5_paralogues) = c("Gene", "Paralogue", "number of predictions")
write.table(top_5_paralogues, file = "data/clinvar/Top_5_genes_with_number_of_annotations_noQC.csv", na="", row.names = FALSE, col.names = TRUE, sep = ",", quote=FALSE)

#carrying on finding out what genes are in the variants being annotated and not for the other parts of the confusion matrixes
var_in_genes_unique_var = data.frame(table(b.paralogous_var_align$Unique_variant_gene_annotations$Gene))
b.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
b.var_in_genes_unique_var = b.var_in_genes_unique_var[b.var_in_genes_unique_var$Var1 %in% clinvar_B_LB_tableized_genes,]
dim(b.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(b.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
b.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
b.not_var_in_genes_unique_var = b.not_var_in_genes_unique_var[b.not_var_in_genes_unique_var$Var1 %in% clinvar_B_LB_tableized_genes,]
length(intersect(b.var_in_genes_unique_var$Var1, b.not_var_in_genes_unique_var$Var1))

var_in_genes_unique_var = data.frame(table(ptob.paralogous_var_align$Unique_variant_gene_annotations$Gene))
ptob.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
dim(ptob.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(ptob.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
ptob.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
length(intersect(ptob.var_in_genes_unique_var$Var1, ptob.not_var_in_genes_unique_var$Var1))

var_in_genes_unique_var = data.frame(table(btob.paralogous_var_align$Unique_variant_gene_annotations$Gene))
btob.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
dim(btob.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(btob.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
btob.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
length(intersect(btob.var_in_genes_unique_var$Var1, btob.not_var_in_genes_unique_var$Var1))

#para_con
p2.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b2.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
ptob2.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")
btob2.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")

conf2 = conf_matrix(p2.paralogous_var_align$true_num_of_paralog_anno, p2.paralogous_var_align$input_variants, b2.paralogous_var_align$true_num_of_paralog_anno, b2.paralogous_var_align$input_variants)
conf_b2 = conf_matrix_benign(ptob2.paralogous_var_align$true_num_of_paralog_anno, ptob2.paralogous_var_align$input_variants, btob2.paralogous_var_align$true_num_of_paralog_anno, btob2.paralogous_var_align$input_variants)
genes_after_QC1 = length(unique(p2.paralogous_var_align$Total_paralog_annotations$Gene))
write.table(unique(p2.paralogous_var_align$Total_paralog_annotations$Gene), file = "data/clinvar/genes_after_QC1.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)
genes_with_PA_QC1 = length(unique(p2.paralogous_var_align$Total_paralog_annotations$SYMBOL))
write.table(unique(p2.paralogous_var_align$Total_paralog_annotations$SYMBOL), file = "data/clinvar/genes_with_PA_QC1.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)

var_removed = var_rem_matrix(conf$con_table, conf2$con_table, p2.paralogous_var_align$paralog_data, b2.paralogous_var_align$paralog_data)
var_removed_b = var_rem_matrix(conf_b$con_table, conf_b2$con_table, ptob2.paralogous_var_align$paralog_data, btob2.paralogous_var_align$paralog_data)

#finding out how many genes in para_con
var_in_genes_unique_var = data.frame(table(p2.paralogous_var_align$Unique_variant_gene_annotations$Gene))
p2.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
dim(p2.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(p2.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
p2.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
length(intersect(p2.var_in_genes_unique_var$Var1, p2.not_var_in_genes_unique_var$Var1))

var_in_genes_unique_var = data.frame(table(b2.paralogous_var_align$Unique_variant_gene_annotations$Gene))
b2.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
dim(b2.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(b2.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
b2.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
length(intersect(b2.var_in_genes_unique_var$Var1, b2.not_var_in_genes_unique_var$Var1))

var_in_genes_unique_var = data.frame(table(ptob2.paralogous_var_align$Unique_variant_gene_annotations$Gene))
ptob2.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
dim(ptob2.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(ptob2.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
ptob2.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
length(intersect(ptob2.var_in_genes_unique_var$Var1, ptob2.not_var_in_genes_unique_var$Var1))

var_in_genes_unique_var = data.frame(table(btob2.paralogous_var_align$Unique_variant_gene_annotations$Gene))
btob2.var_in_genes_unique_var = arrange(var_in_genes_unique_var, desc(Freq))
dim(btob2.paralogous_var_align$Unique_variants_not_annotated)[1]
not_var_in_genes_unique_var = data.frame(table(btob2.paralogous_var_align$Unique_variant_genes_not_annotated$Gene))
btob2.not_var_in_genes_unique_var = arrange(not_var_in_genes_unique_var, desc(Freq))
length(intersect(btob2.var_in_genes_unique_var$Var1, btob2.not_var_in_genes_unique_var$Var1))

#all_con
p3.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b3.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
ptob3.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")
btob3.paralogous_var_align = Paralogous_var_align("data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")

conf3 = conf_matrix(p3.paralogous_var_align$true_num_of_paralog_anno, p3.paralogous_var_align$input_variants, b3.paralogous_var_align$true_num_of_paralog_anno, b3.paralogous_var_align$input_variants)
conf_b3 = conf_matrix_benign(ptob3.paralogous_var_align$true_num_of_paralog_anno, ptob3.paralogous_var_align$input_variants, btob3.paralogous_var_align$true_num_of_paralog_anno, btob3.paralogous_var_align$input_variants)
genes_after_QC2 = length(unique(p3.paralogous_var_align$Total_paralog_annotations$Gene))
write.table(unique(p3.paralogous_var_align$Total_paralog_annotations$Gene), file = "data/clinvar/genes_after_QC2.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)
genes_with_PA_QC2 = length(unique(p3.paralogous_var_align$Total_paralog_annotations$SYMBOL))
write.table(unique(p3.paralogous_var_align$Total_paralog_annotations$SYMBOL), file = "data/clinvar/genes_with_PA_QC2.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)

var_removed2 = var_rem_matrix(conf2$con_table, conf3$con_table, p3.paralogous_var_align$paralog_data, b3.paralogous_var_align$paralog_data)
var_removed_b2 = var_rem_matrix(conf_b2$con_table, conf_b3$con_table, ptob3.paralogous_var_align$paralog_data, btob3.paralogous_var_align$paralog_data)

#alt_con
#pathogenic to pathogenic
p4.Total_paralog_annotations = p3.paralogous_var_align$Total_paralog_annotations[p3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==p3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
# p4.Total_paralog_annotations = p4.Total_paralog_annotations[!duplicated(p4.Total_paralog_annotations$ID.x),]
p4.Total_paralog_annotations = distinct(p4.Total_paralog_annotations)

p4.num_of_paralog_anno = sum(!is.na(p4.Total_paralog_annotations$ID.y))
p4.Unique_variant_annotations = p4.Total_paralog_annotations[!is.na(p4.Total_paralog_annotations$ID.y),]
p4.Unique_variant_annotations = distinct(p4.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p4.true_num_of_paralog_anno = length(p4.Unique_variant_annotations$Variant_pos)

#benign to pathogenic
b4.Total_paralog_annotations = b3.paralogous_var_align$Total_paralog_annotations[b3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==b3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
# b4.Total_paralog_annotations = b4.Total_paralog_annotations[!duplicated(b4.Total_paralog_annotations$ID.x),]
b4.Total_paralog_annotations = distinct(b4.Total_paralog_annotations)

b4.num_of_paralog_anno = sum(!is.na(b4.Total_paralog_annotations$ID.y))
b4.Unique_variant_annotations = b4.Total_paralog_annotations[!is.na(b4.Total_paralog_annotations$ID.y),]
b4.Unique_variant_annotations = distinct(b4.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b4.true_num_of_paralog_anno = length(b4.Unique_variant_annotations$Variant_pos)

#pathogenic to benign
ptob4.Total_paralog_annotations = ptob3.paralogous_var_align$Total_paralog_annotations[ptob3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==ptob3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
ptob4.Total_paralog_annotations = distinct(ptob4.Total_paralog_annotations)

ptob4.num_of_paralog_anno = sum(!is.na(ptob4.Total_paralog_annotations$ID.y))
ptob4.Unique_variant_annotations = ptob4.Total_paralog_annotations[!is.na(ptob4.Total_paralog_annotations$ID.y),]
ptob4.Unique_variant_annotations = distinct(ptob4.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
ptob4.true_num_of_paralog_anno = length(ptob4.Unique_variant_annotations$Variant_pos)

#benign to benign
btob4.Total_paralog_annotations = btob3.paralogous_var_align$Total_paralog_annotations[btob3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==btob3.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
btob4.Total_paralog_annotations = distinct(btob4.Total_paralog_annotations)

btob4.num_of_paralog_anno = sum(!is.na(btob4.Total_paralog_annotations$ID.y))
btob4.Unique_variant_annotations = btob4.Total_paralog_annotations[!is.na(btob4.Total_paralog_annotations$ID.y),]
btob4.Unique_variant_annotations = distinct(btob4.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
btob4.true_num_of_paralog_anno = length(btob4.Unique_variant_annotations$Variant_pos)


conf4 = conf_matrix(p4.true_num_of_paralog_anno, p3.paralogous_var_align$input_variants, b4.true_num_of_paralog_anno, b3.paralogous_var_align$input_variants)
conf_b4 = conf_matrix_benign(ptob4.true_num_of_paralog_anno, ptob3.paralogous_var_align$input_variants, btob4.true_num_of_paralog_anno, btob3.paralogous_var_align$input_variants)
genes_after_QC3 = length(unique(p4.Total_paralog_annotations$Gene))
write.table(unique(p4.Total_paralog_annotations$Gene), file = "data/clinvar/genes_after_QC3.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)
genes_with_PA_QC3 = length(unique(p4.Total_paralog_annotations$SYMBOL))
write.table(unique(p4.Total_paralog_annotations$SYMBOL), file = "data/clinvar/genes_with_PA_QC3.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)

var_removed3 = var_rem_matrix(conf3$con_table, conf4$con_table, p3.paralogous_var_align$paralog_data, b3.paralogous_var_align$paralog_data)
var_removed_b3 = var_rem_matrix(conf_b3$con_table, conf_b4$con_table, ptob3.paralogous_var_align$paralog_data, btob3.paralogous_var_align$paralog_data)

#benign test for only alt con
ptob5.Total_paralog_annotations = ptob.paralogous_var_align$Total_paralog_annotations[ptob.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==ptob.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
ptob5.Total_paralog_annotations = distinct(ptob5.Total_paralog_annotations)

ptob5.num_of_paralog_anno = sum(!is.na(ptob5.Total_paralog_annotations$ID.y))
ptob5.Unique_variant_annotations = ptob5.Total_paralog_annotations[!is.na(ptob5.Total_paralog_annotations$ID.y),]
ptob5.Unique_variant_annotations = distinct(ptob5.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
ptob5.true_num_of_paralog_anno = length(ptob5.Unique_variant_annotations$Variant_pos)

btob5.Total_paralog_annotations = btob.paralogous_var_align$Total_paralog_annotations[btob.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==btob.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
btob5.Total_paralog_annotations = distinct(btob5.Total_paralog_annotations)

btob5.num_of_paralog_anno = sum(!is.na(btob5.Total_paralog_annotations$ID.y))
btob5.Unique_variant_annotations = btob5.Total_paralog_annotations[!is.na(btob5.Total_paralog_annotations$ID.y),]
btob5.Unique_variant_annotations = distinct(btob5.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
btob5.true_num_of_paralog_anno = length(btob5.Unique_variant_annotations$Variant_pos)

conf_b5 = conf_matrix_benign(ptob5.true_num_of_paralog_anno, ptob.paralogous_var_align$input_variants, btob5.true_num_of_paralog_anno, btob.paralogous_var_align$input_variants)

####
#alt_con with para_con but without all_con
#pathogenic to pathogenic
p6.Total_paralog_annotations = p2.paralogous_var_align$Total_paralog_annotations[p2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==p2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
p6.Total_paralog_annotations = distinct(p6.Total_paralog_annotations)

p6.num_of_paralog_anno = sum(!is.na(p6.Total_paralog_annotations$ID.y))
p6.Unique_variant_annotations = p6.Total_paralog_annotations[!is.na(p6.Total_paralog_annotations$ID.y),]
p6.Unique_variant_annotations = distinct(p6.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p6.true_num_of_paralog_anno = length(p6.Unique_variant_annotations$Variant_pos)

#benign to pathogenic
b6.Total_paralog_annotations = b2.paralogous_var_align$Total_paralog_annotations[b2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==b2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
b6.Total_paralog_annotations = distinct(b6.Total_paralog_annotations)

b6.num_of_paralog_anno = sum(!is.na(b6.Total_paralog_annotations$ID.y))
b6.Unique_variant_annotations = b6.Total_paralog_annotations[!is.na(b6.Total_paralog_annotations$ID.y),]
b6.Unique_variant_annotations = distinct(b6.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b6.true_num_of_paralog_anno = length(b6.Unique_variant_annotations$Variant_pos)

#pathogenic to benign
ptob6.Total_paralog_annotations = ptob2.paralogous_var_align$Total_paralog_annotations[ptob2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==ptob2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
ptob6.Total_paralog_annotations = distinct(ptob6.Total_paralog_annotations)

ptob6.num_of_paralog_anno = sum(!is.na(ptob6.Total_paralog_annotations$ID.y))
ptob6.Unique_variant_annotations = ptob6.Total_paralog_annotations[!is.na(ptob6.Total_paralog_annotations$ID.y),]
ptob6.Unique_variant_annotations = distinct(ptob6.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
ptob6.true_num_of_paralog_anno = length(ptob6.Unique_variant_annotations$Variant_pos)

#benign to benign
btob6.Total_paralog_annotations = btob2.paralogous_var_align$Total_paralog_annotations[btob2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.x==btob2.paralogous_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
btob6.Total_paralog_annotations = distinct(btob6.Total_paralog_annotations)

btob6.num_of_paralog_anno = sum(!is.na(btob6.Total_paralog_annotations$ID.y))
btob6.Unique_variant_annotations = btob6.Total_paralog_annotations[!is.na(btob6.Total_paralog_annotations$ID.y),]
btob6.Unique_variant_annotations = distinct(btob6.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
btob6.true_num_of_paralog_anno = length(btob6.Unique_variant_annotations$Variant_pos)


conf6 = conf_matrix(p6.true_num_of_paralog_anno, p2.paralogous_var_align$input_variants, b6.true_num_of_paralog_anno, b2.paralogous_var_align$input_variants)
conf_b6 = conf_matrix_benign(ptob6.true_num_of_paralog_anno, ptob2.paralogous_var_align$input_variants, btob6.true_num_of_paralog_anno, btob2.paralogous_var_align$input_variants)
# genes_after_QC3_w_QC1_no_QC2 = length(unique(p6.Total_paralog_annotations$Gene))
# write.table(unique(p6.Total_paralog_annotations$Gene), file = "data/clinvar/genes_after_QC3_w_QC1_no_QC2.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)
# genes_with_PA_QC3_w_QC1_no_QC2 = length(unique(p6.Total_paralog_annotations$SYMBOL))
# write.table(unique(p6.Total_paralog_annotations$SYMBOL), file = "data/clinvar/genes_with_PA_QC3_w_QC1_no_QC2.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)

var_removed4 = var_rem_matrix(conf2$con_table, conf6$con_table, p2.paralogous_var_align$paralog_data, b2.paralogous_var_align$paralog_data)
var_removed_b4 = var_rem_matrix(conf_b2$con_table, conf_b6$con_table, ptob2.paralogous_var_align$paralog_data, btob2.paralogous_var_align$paralog_data)

####

#Overall_table for predicting pathogenic variants
Variant = c("Pathogenic variants","Benign variants","PPV","Sensitivity","Specificity","P value")
Total = c(dim(p.paralogous_var_align$input_variants)[1],dim(b.paralogous_var_align$input_variants)[1],"NA","NA","NA","NA")
Paralogue_Annotation_no_QC = c(format(conf$con_table[2], scientific = FALSE),format(conf$con_table[4], scientific = FALSE),signif(conf$PPV, digits = 2),signif(conf$Sensitivity, digits = 2),signif(conf$Specificity, digits = 2),signif(conf$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC1 = c(format(conf2$con_table[2], scientific = FALSE),format(conf2$con_table[4], scientific = FALSE),signif(conf2$PPV, digits = 2), signif(conf2$Sensitivity, digits = 2),signif(conf2$Specificity, digits = 2), signif(conf2$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC2 = c(format(conf3$con_table[2], scientific = FALSE),format(conf3$con_table[4], scientific = FALSE),signif(conf3$PPV, digits = 2),signif(conf3$Sensitivity, digits = 2),signif(conf3$Specificity, digits = 2),signif(conf3$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC3 = c(format(conf4$con_table[2], scientific = FALSE),format(conf4$con_table[4], scientific = FALSE),signif(conf4$PPV, digits = 2),signif(conf4$Sensitivity, digits = 2),signif(conf4$Specificity, digits = 2),signif(conf4$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC3_w_QC1_no_QC2 = c(format(conf6$con_table[2], scientific = FALSE),format(conf6$con_table[4], scientific = FALSE),signif(conf6$PPV, digits = 2),signif(conf6$Sensitivity, digits = 2),signif(conf6$Specificity, digits = 2),signif(conf6$Pvalue$p.value, digits = 2))

Variants_removed_after_PA_QC1 = c(format(var_removed$con_table[2], scientific = FALSE),format(var_removed$con_table[4], scientific = FALSE),signif(var_removed$PPV, digits = 2),signif(var_removed$Sensitivity, digits = 2),signif(var_removed$Specificity, digits = 2),signif(var_removed$Pvalue$p.value, digits = 2))
Variants_removed_after_PA_QC2 = c(format(var_removed2$con_table[2], scientific = FALSE),format(var_removed2$con_table[4], scientific = FALSE),signif(var_removed2$PPV, digits = 2),signif(var_removed2$Sensitivity, digits = 2),signif(var_removed2$Specificity, digits = 2),signif(var_removed2$Pvalue$p.value, digits = 2))
Variants_removed_after_PA_QC3 = c(format(var_removed3$con_table[2], scientific = FALSE),format(var_removed3$con_table[4], scientific = FALSE),signif(var_removed3$PPV, digits = 2),signif(var_removed3$Sensitivity, digits = 2),signif(var_removed3$Specificity, digits = 2),signif(var_removed3$Pvalue$p.value, digits = 2))
Variants_removed_after_PA_QC3_w_QC1_no_QC2 = c(format(var_removed4$con_table[2], scientific = FALSE),format(var_removed4$con_table[4], scientific = FALSE),signif(var_removed4$PPV, digits = 2),signif(var_removed4$Sensitivity, digits = 2),signif(var_removed4$Specificity, digits = 2),signif(var_removed4$Pvalue$p.value, digits = 2))

# Paralogue_Annotation_no_QC = c(format(conf$con_table[2], scientific = FALSE),format(conf$con_table[4], scientific = FALSE),conf$PPV,conf$Sensitivity,conf$Pvalue$p.value)
# Variants_remaining_after_PA_QC1 = c(conf2$con_table[2],conf2$con_table[4],conf2$PPV, conf2$Sensitivity, conf2$Pvalue$p.value)
# 
# Variants_removed_after_PA_QC1 = c(var_removed$con_table[2],var_removed$con_table[4],var_removed$PPV,"NA",var_removed$Pvalue$p.value)
# Variants_remaining_after_PA_QC2 = c(conf3$con_table[2],conf3$con_table[4],conf3$PPV,conf3$Sensitivity,conf3$Pvalue$p.value)
# Variants_removed_after_PA_QC2 = c(var_removed2$con_table[2],var_removed2$con_table[4],var_removed2$PPV,"NA",var_removed2$Pvalue$p.value)
# Variants_remaining_after_PA_QC3 = c(conf4$con_table[2],conf4$con_table[4],conf4$PPV,conf4$Sensitivity,conf4$Pvalue$p.value)
# Variants_removed_after_PA_QC3 = c(var_removed3$con_table[2],var_removed3$con_table[4],var_removed3$PPV,"NA",var_removed3$Pvalue$p.value)

clinvar_set = data.frame(#Variant,
                         Total,
                         Paralogue_Annotation_no_QC,
                         Variants_remaining_after_PA_QC1,
                         Variants_remaining_after_PA_QC2,
                         Variants_remaining_after_PA_QC3,
                         Variants_remaining_after_PA_QC3_w_QC1_no_QC2,
                         Variants_removed_after_PA_QC1,
                         Variants_removed_after_PA_QC2,
                         Variants_removed_after_PA_QC3,
                         Variants_removed_after_PA_QC3_w_QC1_no_QC2
                         )

clinvar_set = t(clinvar_set)
colnames(clinvar_set) = c("Pathogenic variants","Benign variants","PPV","Sensitivity","Specificity","P value")
rownames(clinvar_set) = c("Total variants", 
                          "Paralogue Annotation", 
                          "Variants remaining after QC1",
                          "Variants remaining after QC2", 
                          "Variants remaining after QC3", 
                          "Variants remaining after QC3, no QC2", 
                          "Variants removed from QC1", 
                          "Variants removed from QC2",  
                          "Variants removed from QC3",
                          "Variants removed from QC3, no QC2")
write.table(clinvar_set, file = "data/clinvar/clinvar_set.csv", na="", row.names = TRUE, col.names = TRUE, sep = ",", quote=FALSE)
kable(clinvar_set)
# pander(clinvar_set2)
# pander(clinvar_set3)

clinvar_set_genes = data.frame(c("Total","Genes with Paralogue Annotations - no QC", "Genes remaining after QC1", "Genes with Paralogue Annotations - QC1", "Genes remaining after QC2", "Genes with Paralogue Annotations - QC2", "Genes remaining after QC3", "Genes with Paralogue Annotations - QC3"),
                               c(genes_before_PA,genes_with_PA_noQC,genes_after_QC1,genes_with_PA_QC1,genes_after_QC2,genes_with_PA_QC2,genes_after_QC3,genes_with_PA_QC3))
colnames(clinvar_set_genes) = c("","Number of genes")
kable(clinvar_set_genes)

#Overall_table for predicting Benign variants
Variant = c("Pathogenic variants","Benign variants","PPV","Sensitivity","Specificity","P value")
Total = c(dim(p.paralogous_var_align$input_variants)[1],dim(b.paralogous_var_align$input)[1],"NA","NA","NA","NA")
Paralogue_Annotation_no_QC = c(format(conf_b$con_table[2], scientific = FALSE),format(conf_b$con_table[4], scientific = FALSE),signif(conf_b$PPV, digits = 2),signif(conf_b$Sensitivity, digits = 2),signif(conf_b$Specificity, digits = 2),signif(conf_b$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC1 = c(format(conf_b2$con_table[2], scientific = FALSE),format(conf_b2$con_table[4], scientific = FALSE),signif(conf_b2$PPV, digits = 2), signif(conf_b2$Sensitivity, digits = 2),signif(conf_b2$Specificity, digits = 2), signif(conf_b2$Pvalue$p.value, digits = 2))
Variants_removed_after_PA_QC1 = c(format(var_removed_b$con_table[2], scientific = FALSE),format(var_removed_b$con_table[4], scientific = FALSE),signif(var_removed_b$PPV, digits = 2),signif(var_removed_b$Sensitivity, digits = 2),signif(var_removed_b$Specificity, digits = 2),signif(var_removed_b$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC2 = c(format(conf_b3$con_table[2], scientific = FALSE),format(conf_b3$con_table[4], scientific = FALSE),signif(conf_b3$PPV, digits = 2),signif(conf_b3$Sensitivity, digits = 2),signif(conf_b3$Specificity, digits = 2),signif(conf_b3$Pvalue$p.value, digits = 2))
Variants_removed_after_PA_QC2 = c(format(var_removed_b2$con_table[2], scientific = FALSE),format(var_removed_b2$con_table[4], scientific = FALSE),signif(var_removed_b2$PPV, digits = 2),signif(var_removed_b2$Sensitivity, digits = 2),signif(var_removed_b2$Specificity, digits = 2),signif(var_removed_b2$Pvalue$p.value, digits = 2))
Variants_remaining_after_PA_QC3 = c(format(conf_b4$con_table[2], scientific = FALSE),format(conf_b4$con_table[4], scientific = FALSE),signif(conf_b4$PPV, digits = 2),signif(conf_b4$Sensitivity, digits = 2),signif(conf_b4$Specificity, digits = 2),signif(conf_b4$Pvalue$p.value, digits = 2))
Variants_removed_after_PA_QC3 = c(format(var_removed_b3$con_table[2], scientific = FALSE),format(var_removed_b3$con_table[4], scientific = FALSE),signif(var_removed_b3$PPV, digits = 2),signif(var_removed_b3$Sensitivity, digits = 2),signif(var_removed_b3$Specificity, digits = 2),signif(var_removed_b3$Pvalue$p.value, digits = 2))

# Paralogue_Annotation_no_QC = c(format(conf$con_table[2], scientific = FALSE),format(conf$con_table[4], scientific = FALSE),conf$PPV,conf$Sensitivity,conf$Pvalue$p.value)
# Variants_remaining_after_PA_QC1 = c(conf2$con_table[2],conf2$con_table[4],conf2$PPV, conf2$Sensitivity, conf2$Pvalue$p.value)
# 
# Variants_removed_after_PA_QC1 = c(var_removed$con_table[2],var_removed$con_table[4],var_removed$PPV,"NA",var_removed$Pvalue$p.value)
# Variants_remaining_after_PA_QC2 = c(conf3$con_table[2],conf3$con_table[4],conf3$PPV,conf3$Sensitivity,conf3$Pvalue$p.value)
# Variants_removed_after_PA_QC2 = c(var_removed2$con_table[2],var_removed2$con_table[4],var_removed2$PPV,"NA",var_removed2$Pvalue$p.value)
# Variants_remaining_after_PA_QC3 = c(conf4$con_table[2],conf4$con_table[4],conf4$PPV,conf4$Sensitivity,conf4$Pvalue$p.value)
# Variants_removed_after_PA_QC3 = c(var_removed3$con_table[2],var_removed3$con_table[4],var_removed3$PPV,"NA",var_removed3$Pvalue$p.value)

clinvar_set_benign = data.frame(#Variant,
                         Total,
                         Paralogue_Annotation_no_QC,
                         Variants_remaining_after_PA_QC1,
                         Variants_remaining_after_PA_QC2,
                         Variants_remaining_after_PA_QC3,
                         Variants_removed_after_PA_QC1,
                         Variants_removed_after_PA_QC2,
                         Variants_removed_after_PA_QC3
                         )

clinvar_set_benign = t(clinvar_set_benign)
colnames(clinvar_set_benign) = c("Pathogenic variants","Benign variants","PPV","Sensitivity","Specificity","P value")
rownames(clinvar_set_benign) = c("Total variants", "Paralogue Annotation", "Variants remaining after QC1", "Variants remaining after QC2", "Variants remaining after QC3", "Variants removed from QC1", "Variants removed from QC2", "Variants removed from QC3")
write.table(clinvar_set_benign, file = "data/clinvar/clinvar_set_benign.csv", na="", row.names = TRUE, col.names = TRUE, sep = ",", quote=FALSE)
kable(clinvar_set_benign)
```

Prior to annotation, there were `r length(clinvar_P_LP_tableized[!is.na(clinvar_P_LP_tableized$Paralogue_Vars),1])` variants that aligned to at least one paralogous equivalent position according to ensembl (`r length(clinvar_P_LP_tableized[!is.na(clinvar_P_LP_tableized$Paralogue_Vars),1])/length(clinvar_P_LP_tableized$CHROM)`)

The full analysis of clinvar variants is shown in **table \ref{fig:clinvar_table}**. In summary, `r conf$con_table[1]` Pathogenic and Likely Pathogenic (P/LP) variants and `r conf$con_table[2]` Benign and Likey Benign (B/LB) variants from clinvar had paralogue annotations. With no quality control, `r conf$con_table[2]` known P/LP  and `r conf$con_table[4]` known B/LB variants were predicted to be pathogenic, given a PPV and sensitivity of `r conf$PPV` and `r conf$Sensitivity` respectively. Comparatively, predicting benign variants was not as reliable. With `r conf_b$con_table[2]` known P/LP and `r conf_b$con_table[4]` known B/LB variants predicted to be benign. Though the proportional differene is statistically significant with a p-value of $`r conf_b$Pvalue$p.value`$, this lead to a PPV and sensitivity of `r conf_b$PPV` and `r conf_b$Sensitivity` respectively. 

Using the aforementioned quality control steps to increase the stringency of conservation across alignment columns in regards to reference and alternate amino acid alleles shows improvement to PPV and decrease in sensitivity over all for predicting pathogenic variants. But this does not help the case for predicting benign variants. The PPV does not improve significantly to a reliable level. Therefore, it can be concluded that at least with the dataset used in this study, paralogue annotation can be used as a variant classification method for predicting pathogenic variants, but not benign.


Looking at pipeline of study:
```{r, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE}
x = read.csv(file="/media/nick/821ED5711ED55F2B/Users/Nick/Documents/PhD/Paralogues/data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc", sep = "\t", header=0, stringsAsFactors = 0, fill=FALSE, quote="", na.strings=".", comment.char="#")
y = read.csv(file="/media/nick/821ED5711ED55F2B/Users/Nick/Documents/PhD/Paralogues/data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.vcf", sep = "\t", header=0, stringsAsFactors = 0, fill=FALSE, quote="", na.strings=".", comment.char="#")
length(unique(x$V3)) #short one
length(unique(y$V3)) #longer one
setdiff(unique(y$V3),unique(x$V3))

x1 = read.csv(file="/media/nick/821ED5711ED55F2B/Users/Nick/Documents/PhD/Paralogues/data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc", sep = "\t", header=0, stringsAsFactors = 0, fill=FALSE, quote="", na.strings=".", comment.char="#")
y1 = read.csv(file="/media/nick/821ED5711ED55F2B/Users/Nick/Documents/PhD/Paralogues/data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.vcf", sep = "\t", header=0, stringsAsFactors = 0, fill=FALSE, quote="", na.strings=".", comment.char="#")
setdiff(unique(y1$V3),unique(x1$V3))
```
NB that in P/LP 11 variants with IDs: 16549; 525690; 12282; 3882; 29695; 13538; 7226; 17114; 7366; 4576; and 504370 are actually deletions so VEP does not predict anything and does not output them in results. Hence out_paraloc file has 11 less lines than the vcf file.
For B/LB theres 295 dels.

```{r, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE}
x = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs", sep = "\t", header=0, stringsAsFactors = 0, fill=FALSE, quote="", na.strings=".", comment.char="#")
x0 = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", sep="\t", header=FALSE, col.names = c("Variant_pos", "ID", "Gene", "REF", "ALT", paste("paralog", 1:147, sep = "")), quote="", na.strings=".", comment.char="#", stringsAsFactors = 0)
x = distinct(x0[c("Variant_pos", "ID", "REF", "ALT")])
library(plyr)
y = ddply(x0[,c("Variant_pos", "ID", "REF", "ALT")],.(Variant_pos, ID, REF, ALT),nrow)
y = y[y$V1>1,]
dim(y)
z = x0[x0$ID %in% y,]

b = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs", sep = "\t", header=0, stringsAsFactors = 0, fill=FALSE, quote="", na.strings=".", comment.char="#")
b0 = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", sep="\t", header=FALSE, col.names = c("Variant_pos", "ID", "Gene", "REF", "ALT", paste("paralog", 1:147, sep = "")), quote="", na.strings=".", comment.char="#", stringsAsFactors = 0)
b = distinct(b0[c("Variant_pos", "ID", "REF", "ALT")])
c = ddply(b0[,c("Variant_pos", "ID", "REF", "ALT")],.(Variant_pos, ID, REF, ALT),nrow)
c = c[c$V1>1,]
dim(c)
```
NB there's 435 unique variants lying in overlapping genes in the P/LP. Overall there's 489 instances of variants in a position that covers two genes (i.e. one instance of two gene overlapp)
```{r, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE}
x = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", sep="\t", header=FALSE, col.names = c("Variant_pos", "ID", "Gene", "REF", "ALT", paste("paralog", 1:147, sep = "")), quote="", na.strings=".", comment.char="#", stringsAsFactors = 0)
setdiff(unique(clinvar_P_LP_genes_w_paralogs),unique(x$Gene))
setdiff(unique(x$Gene),unique(clinvar_P_LP_genes_w_paralogs))
# write.table(gene_EFs, file = "data/case_and_control/gene_EFs.csv", na="", row.names = FALSE, col.names = TRUE, sep = ",", quote=FALSE)
length(intersect(unique(x$Gene),unique(clinvar_P_LP_tableized_genes)))
write.table(setdiff(unique(clinvar_P_LP_tableized_genes),unique(x$Gene)), file = "data/clinvar/genes_in_out_paraloc_paralogs2.noQC.txt", row.names = FALSE, col.names = FALSE)

y = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", sep="\t", header=FALSE, col.names = c("Variant_pos", "ID", "Gene", "REF", "ALT", paste("paralog", 1:147, sep = "")), quote="", na.strings=".", comment.char="#", stringsAsFactors = 0)
length(intersect(unique(y$Gene),unique(clinvar_B_LB_tableized_genes)))
```
NB in clinvar_P_LP_genes_w_paralogs there are 3621 genes with paralogues but in the out_paraloc_paralogs file there are only 2451 genes. This is because there 1170 genes in clinvar_P_LP_genes_w_paralogs that are not in out_paraloc_paralogs because though there are variants in these genes they do not align to anything. There are also cases of over-lapping genes and so while one gene is reported the other might not. Also the same for some readthrough transcripts. Several genes with designation "AC######.#" go through non-sense mediated decay.

## Para-Z scores
The filtering steps outlined above take a more binaray approach to taking account the conservativeness of amino acid positions in the alignments. They only consider if amino acids in question share the the same amino acid or not. The Para-Z scores defined by @Lal2017 on the other hand takes a more quantitative approach to this by representing a numeric integer value of how conserved each amino acid position is across the same paralogue family. As shown by **fig.** \ref{fig:paraz}, one can use the Para-Z scores as cutoff thresholds for how paralogue conserved alignments must be before considering annotations. The more stringent the cutoff the lower the false positive rate is, but as expected the lower the sensitivity is also. Regardless both methods validate the concept that the more conserved amino acid positions are when transfering annotation the more likely annotations will be true positives as one would expect.
\newpage

```{r paraz, message=FALSE, cache=TRUE,  results='asis', warning=FALSE, echo=FALSE, fig.width=4,fig.height=3,fig.cap="\\label{fig:paraz}ROC space plot of pathogenic/likely pathogenic clinvar variants predicted as pathogenic by paralogue annotation at varying degrees of Para-Z scores"}
Sensitivity_list = vector()
Sens_CI_list = vector()
FPR_list = vector()
con_table_list = matrix(, ncol = 2)
PPV_list = vector()
PPV_CI_list = vector()
P_value_list = vector()
Specificity_list = vector()
##Table##
paraz_set = t(data.frame(Variant,Total))
roc_data = data.frame()
####
for (i in seq(0,11)){
#noQC
#pathogenic to pathogenic
# p.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
  p.paraz_var_align = Paralogous_var_align(paraz_cutoff = i,"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")

#benign to pathogenic
# b.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
  b.paraz_var_align = Paralogous_var_align(paraz_cutoff = i,"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")

  paraz.conf = conf_matrix(p.paraz_var_align$true_num_of_paralog_anno, p.paraz_var_align$input_variants, b.paraz_var_align$true_num_of_paralog_anno, b.paraz_var_align$input_variants)
  Sensitivity_list = c(Sensitivity_list, paraz.conf$Sensitivity)
  Sens_CI_list = c(Sens_CI_list, 1.96*paraz.conf$Sensitivity_SE)
  FPR_list = c(FPR_list, paraz.conf$FPR)
  con_table_list = rbind(con_table_list, paraz.conf$con_table)
  PPV_list = c(PPV_list, paraz.conf$PPV)
  PPV_CI_list= c(PPV_CI_list, 1.96*paraz.conf$PPV_SE)
  P_value_list = c(P_value_list, paraz.conf$Pvalue$p.value)
  Specificity_list = c(Specificity_list, paraz.conf$Specificity)
  
  PA_threshold_row = c(format(paraz.conf$con_table[2], scientific = FALSE), format(paraz.conf$con_table[2,2], scientific = FALSE), round(paraz.conf$PPV, digits = 3), round(paraz.conf$Sensitivity, digits = 3), round(paraz.conf$Specificity,digits = 3), signif(paraz.conf$Pvalue$p.value, digits = 3))
  paraz_set = rbind(paraz_set, PA_threshold_row)
}
##Graph
# roc_data = do.call(rbind, Map(data.frame, Sensitivity = Sensitivity_list, FPR = FPR_list, cutoff = seq(0,11)))
roc_data = do.call(rbind, Map(data.frame, Sensitivity = Sensitivity_list, FPR = FPR_list))
roc_data$Type = "Para_z"

# colnames(roc_data) = c("Sensitivity", "FPR", "type")
# roc_data$type="Para_z"
roc_data = rbind(roc_data, c(as.numeric(conf$Sensitivity),as.numeric(conf$FPR)))
roc_data[13,3]="noQC"
roc_data = rbind(roc_data, c(as.numeric(conf2$Sensitivity),as.numeric(conf2$FPR)))
roc_data[14,3]="QC1"
roc_data = rbind(roc_data, c(as.numeric(conf3$Sensitivity),as.numeric(conf3$FPR)))
roc_data[15,3]="QC2"
roc_data = rbind(roc_data, c(as.numeric(conf4$Sensitivity),as.numeric(conf4$FPR)))
roc_data[16,3]="QC3"

# gg = ggplot(roc_data, aes(x=FPR, y=Sensitivity, colour=cutoff)) + geom_point() + xlim(0,1) + ylim(0,1)
# gg = gg + geom_abline(intercept = 0, slope = 1, linetype = "dotted")
# 
# ggplot(roc_data, aes(x=FPR, y=Sensitivity, colour = type)) + geom_point() + xlim(0,0.05) + ylim(0,0.25) + scale_color_gradient(low="blue", high="red", limits=c(0, 11))
gg = ggplot(roc_data, aes(x=FPR, y=Sensitivity, colour = Type)) + geom_point(size=2) + xlim(0,0.05) + ylim(0,0.25) + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14, 
        hjust = 0.25)) +labs(title = "Receiver operating characteristic space plot of\nParalogue Annotation") + theme(axis.text = element_text(size = 13))
ggsave("images/ROC_plot.png")

# prc_data = do.call(rbind, Map(data.frame, Precision = PPV_list, Sensitivity = Sensitivity_list, cutoff = seq(0,11)))
prc_data = do.call(rbind, Map(data.frame, Precision = PPV_list, Precision_CI = PPV_CI_list, Sensitivity = Sensitivity_list, Sensitivity_CI = Sens_CI_list))
prc_data$Type = "Para_z"

prc_data = rbind(prc_data, c(as.numeric(conf$PPV),as.numeric(1.96*conf$PPV_SE),as.numeric(conf$Sensitivity),as.numeric(1.96*conf$Sensitivity_SE)))
prc_data[13,5]="noQC"
prc_data = rbind(prc_data, c(as.numeric(conf2$PPV),as.numeric(1.96*conf2$PPV_SE),as.numeric(conf2$Sensitivity),as.numeric(1.96*conf2$Sensitivity_SE)))
prc_data[14,5]="QC1"
prc_data = rbind(prc_data, c(as.numeric(conf3$PPV),as.numeric(1.96*conf3$PPV_SE),as.numeric(conf3$Sensitivity),as.numeric(1.96*conf3$Sensitivity_SE)))
prc_data[15,5]="QC2"
prc_data = rbind(prc_data, c(as.numeric(conf4$PPV),as.numeric(1.96*conf4$PPV_SE),as.numeric(conf4$Sensitivity),as.numeric(1.96*conf$Sensitivity_SE)))
prc_data[16,5]="QC3"

# gg2 = ggplot(prc_data, aes(x=Sensitivity, y=Precision, colour=cutoff)) + geom_point() + xlim(0,1) + ylim(0,1)
# gg2 = gg2 + geom_abline(intercept = 0, slope = 1, linetype = "dotted")
# 
# ggplot(prc_data, aes(x=Sensitivity, y=Precision, colour = cutoff)) + geom_point() + xlim(0.1,0.2) + ylim(0.85,1) + scale_color_gradient(low="blue", high="red", limits=c(0, 11))
gg2 = ggplot(prc_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=5) + geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,width=5) +
  scale_shape_manual(values=c(15, 1, 16, 17, 18)) + xlim(0,0.25) + ylim(0.85,1)  + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5))
ggsave("images/PRC_plot.png")

prc_data2 = prc_data[13:16,]
# prc_data2 = rbind(prc_data2, c(0.619,0.032))
# prc_data2[5,3] = "removed by QC1"
# prc_data2 = rbind(prc_data2, c(0.931,0.083))
# prc_data2[6,3] = "removed by QC2"
# prc_data2 = rbind(prc_data2, c(0.981,0.058))
# prc_data2[7,3] = "removed by QC3"
gg2_5 = ggplot(prc_data2, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=.5) +
  scale_shape_manual(values=c(15, 16, 17, 18)) + xlim(0,0.25) + ylim(0.85,1)  + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5))
ggsave("images/PRC_with_removed_plot.png")



benign_data = data.frame(Precision = c(0.391,0.258,0.169,0.963), Precision_CI = c(0.0161531,0.02014389,0.02741531,0.0356178), Sensitivity = c(0.092,0.032,0.008,0.007), Sensitivity_CI = c(0.004658095,0.002811256,0.001446604,0.001341913),Type = c("noQC","QC1","QC2","QC3"))
gg3 = ggplot(benign_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) +
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=0) +
  scale_shape_manual(values=c(15, 16, 17, 18)) + xlim(0,0.25) + ylim(0,1)  + theme(plot.subtitle = element_text(vjust = 1),
                                                                                                                                  plot.caption = element_text(vjust = 1),
                                                                                                                                  axis.title = element_text(size = 16), 
plot.title = element_text(size = 14)) + labs(title = "Precision-Sensitivity space plot of Paralogue Annotation", x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5))
ggsave("images/benign_PRC_plot.png")


kable(paraz_set)
```

Can also apply Paraz scores after QC1

```{r paraz1, message=FALSE, cache=TRUE,  results='asis', warning=FALSE, echo=FALSE, fig.width=4,fig.height=3,fig.cap="\\label{fig:paraz}ROC space plot of pathogenic/likely pathogenic clinvar variants predicted as pathogenic by paralogue annotation at varying degrees of Para-Z scores"}
Sensitivity_list = vector()
Sens_CI_list = vector()
FPR_list = vector()
con_table_list = matrix(, ncol = 2)
PPV_list = vector()
PPV_CI_list = vector()
P_value_list = vector()
Specificity_list = vector()
##Table##
paraz_set = t(data.frame(Variant,Total))
roc_data = data.frame()
####
for (i in seq(0,11)){
#noQC
#pathogenic to pathogenic
# p.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
  p.paraz_var_align = Paralogous_var_align(paraz_cutoff = i,"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")

#benign to pathogenic
# b.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
  b.paraz_var_align = Paralogous_var_align(paraz_cutoff = i,"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")

  paraz.conf = conf_matrix(p.paraz_var_align$true_num_of_paralog_anno, p.paraz_var_align$input_variants, b.paraz_var_align$true_num_of_paralog_anno, b.paraz_var_align$input_variants)
  Sensitivity_list = c(Sensitivity_list, paraz.conf$Sensitivity)
  Sens_CI_list = c(Sens_CI_list, 1.96*paraz.conf$Sensitivity_SE)
  FPR_list = c(FPR_list, paraz.conf$FPR)
  con_table_list = rbind(con_table_list, paraz.conf$con_table)
  PPV_list = c(PPV_list, paraz.conf$PPV)
  PPV_CI_list= c(PPV_CI_list, 1.96*paraz.conf$PPV_SE)
  P_value_list = c(P_value_list, paraz.conf$Pvalue$p.value)
  Specificity_list = c(Specificity_list, paraz.conf$Specificity)
  
  PA_threshold_row = c(format(paraz.conf$con_table[2], scientific = FALSE), format(paraz.conf$con_table[2,2], scientific = FALSE), round(paraz.conf$PPV, digits = 3), round(paraz.conf$Sensitivity, digits = 3), round(paraz.conf$Specificity,digits = 3), signif(paraz.conf$Pvalue$p.value, digits = 3))
  paraz_set = rbind(paraz_set, PA_threshold_row)
}
##Graph
# roc_data = do.call(rbind, Map(data.frame, Sensitivity = Sensitivity_list, FPR = FPR_list, cutoff = seq(0,11)))
roc_data = do.call(rbind, Map(data.frame, Sensitivity = Sensitivity_list, FPR = FPR_list))
roc_data$Type = "Para_z"

# colnames(roc_data) = c("Sensitivity", "FPR", "type")
# roc_data$type="Para_z"
roc_data = rbind(roc_data, c(as.numeric(conf$Sensitivity),as.numeric(conf$FPR)))
roc_data[13,3]="noQC"
roc_data = rbind(roc_data, c(as.numeric(conf2$Sensitivity),as.numeric(conf2$FPR)))
roc_data[14,3]="QC1"
roc_data = rbind(roc_data, c(as.numeric(conf3$Sensitivity),as.numeric(conf3$FPR)))
roc_data[15,3]="QC2"
roc_data = rbind(roc_data, c(as.numeric(conf4$Sensitivity),as.numeric(conf4$FPR)))
roc_data[16,3]="QC3"

# gg = ggplot(roc_data, aes(x=FPR, y=Sensitivity, colour=cutoff)) + geom_point() + xlim(0,1) + ylim(0,1)
# gg = gg + geom_abline(intercept = 0, slope = 1, linetype = "dotted")
# 
# ggplot(roc_data, aes(x=FPR, y=Sensitivity, colour = type)) + geom_point() + xlim(0,0.05) + ylim(0,0.25) + scale_color_gradient(low="blue", high="red", limits=c(0, 11))
gg = ggplot(roc_data, aes(x=FPR, y=Sensitivity, colour = Type)) + geom_point(size=2) + xlim(0,0.05) + ylim(0,0.25) + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14, 
        hjust = 0.25)) +labs(title = "Receiver operating characteristic space plot of\nParalogue Annotation") + theme(axis.text = element_text(size = 13))
ggsave("images/ROC_plot_paraz_with_QC1.png")

# prc_data = do.call(rbind, Map(data.frame, Precision = PPV_list, Sensitivity = Sensitivity_list, cutoff = seq(0,11)))
prc_data = do.call(rbind, Map(data.frame, Precision = PPV_list, Precision_CI = PPV_CI_list, Sensitivity = Sensitivity_list, Sensitivity_CI = Sens_CI_list))
prc_data$Type = "Para_z"

prc_data = rbind(prc_data, c(as.numeric(conf$PPV),as.numeric(1.96*conf$PPV_SE),as.numeric(conf$Sensitivity),as.numeric(1.96*conf$Sensitivity_SE)))
prc_data[13,5]="noQC"
prc_data = rbind(prc_data, c(as.numeric(conf2$PPV),as.numeric(1.96*conf2$PPV_SE),as.numeric(conf2$Sensitivity),as.numeric(1.96*conf2$Sensitivity_SE)))
prc_data[14,5]="QC1"
prc_data = rbind(prc_data, c(as.numeric(conf3$PPV),as.numeric(1.96*conf3$PPV_SE),as.numeric(conf3$Sensitivity),as.numeric(1.96*conf3$Sensitivity_SE)))
prc_data[15,5]="QC2"
prc_data = rbind(prc_data, c(as.numeric(conf4$PPV),as.numeric(1.96*conf4$PPV_SE),as.numeric(conf4$Sensitivity),as.numeric(1.96*conf4$Sensitivity_SE)))
prc_data[16,5]="QC3"
prc_data = rbind(prc_data, c(as.numeric(conf6$PPV),as.numeric(1.96*conf6$PPV_SE),as.numeric(conf6$Sensitivity),as.numeric(1.96*conf6$Sensitivity_SE)))
prc_data[17,5]="QC3 (skipped QC2)"

# gg2 = ggplot(prc_data, aes(x=Sensitivity, y=Precision, colour=cutoff)) + geom_point() + xlim(0,1) + ylim(0,1)
# gg2 = gg2 + geom_abline(intercept = 0, slope = 1, linetype = "dotted")
# 
# ggplot(prc_data, aes(x=Sensitivity, y=Precision, colour = cutoff)) + geom_point() + xlim(0.1,0.2) + ylim(0.85,1) + scale_color_gradient(low="blue", high="red", limits=c(0, 11))
gg2 = ggplot(prc_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=5) + geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,width=5) +
  scale_shape_manual(values=c(15, 1, 16, 17, 18, 19)) + xlim(0,0.25) + ylim(0.85,1)  + theme(plot.subtitle = element_text(vjust = 1), 
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5))
ggsave("images/PRC_plot_paraz_with_QC1_and_QC3_no_QC2.png")



kable(paraz_set)
```

## Case example: lack of paralogue annotation on 8 sarcomeric genes subset and use of etiological fractions as alternate validation 
```{r echo = FALSE, cache=TRUE, results='asis', warning=FALSE}
#noQC
# p.sarco_var_align = Subset_var_align_gene(c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
# b.sarco_var_align = Subset_var_align_gene(c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
# conf = conf_matrix(p.sarco_var_align$num_of_paralog_anno, p.sarco_var_align$paralog_data, b.sarco_var_align$num_of_paralog_anno, b.sarco_var_align$paralog_data)

p.sarco_var_align = Paralogous_var_align(Subset = c("Gene","MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.sarco_var_align = Paralogous_var_align(Subset = c("Gene","MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
conf = conf_matrix(p.sarco_var_align$true_num_of_paralog_anno, p.sarco_var_align$paralog_data, b.sarco_var_align$true_num_of_paralog_anno, b.sarco_var_align$paralog_data)


#para_con
# p2.sarco_var_align = Subset_var_align_gene(c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
# b2.sarco_var_align = Subset_var_align_gene(c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
# conf2 = conf_matrix(p2.sarco_var_align$num_of_paralog_anno, p2.sarco_var_align$paralog_data, b2.sarco_var_align$num_of_paralog_anno, b2.sarco_var_align$paralog_data)
p2.sarco_var_align = Paralogous_var_align(Subset = c("Gene","MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b2.sarco_var_align = Paralogous_var_align(Subset = c("Gene","MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
conf2 = conf_matrix(p2.sarco_var_align$true_num_of_paralog_anno, p2.sarco_var_align$paralog_data, b2.sarco_var_align$true_num_of_paralog_anno, b2.sarco_var_align$paralog_data)

var_removed = var_rem_matrix(conf$con_table, conf2$con_table, p2.sarco_var_align$paralog_data, b2.sarco_var_align$paralog_data)

#all_con
# p3.sarco_var_align = Subset_var_align_gene(c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
# b3.sarco_var_align = Subset_var_align_gene(c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
# conf3 = conf_matrix(p3.sarco_var_align$num_of_paralog_anno, p3.sarco_var_align$paralog_data, b3.sarco_var_align$num_of_paralog_anno, b3.sarco_var_align$paralog_data)
p3.sarco_var_align = Paralogous_var_align(Subset = c("Gene","MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b3.sarco_var_align = Paralogous_var_align(Subset = c("Gene","MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1"),"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
conf3 = conf_matrix(p3.sarco_var_align$true_num_of_paralog_anno, p3.sarco_var_align$paralog_data, b3.sarco_var_align$true_num_of_paralog_anno, b3.sarco_var_align$paralog_data)

var_removed2 = var_rem_matrix(conf2$con_table, conf3$con_table, p3.sarco_var_align$paralog_data, b3.sarco_var_align$paralog_data)

#alt_con
#pathogenic to pathogenic
p4.Total_paralog_annotations = p3.sarco_var_align$Total_paralog_annotations[p3.sarco_var_align$Total_paralog_annotations$ALT_Amino_acids.x==p3.sarco_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
p4.Total_paralog_annotations = distinct(p4.Total_paralog_annotations)
p4.num_of_paralog_anno = sum(!is.na(p4.Total_paralog_annotations$ID.y))
p4.Unique_variant_annotations = p4.Total_paralog_annotations[!is.na(p4.Total_paralog_annotations$ID.y),]
p4.Unique_variant_annotations = distinct(p4.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p4.true_num_of_paralog_anno = length(p4.Unique_variant_annotations$Variant_pos)

#benign to pathogenic
b4.Total_paralog_annotations = b3.sarco_var_align$Total_paralog_annotations[b3.sarco_var_align$Total_paralog_annotations$ALT_Amino_acids.x==b3.sarco_var_align$Total_paralog_annotations$ALT_Amino_acids.y,]
b4.Total_paralog_annotations = distinct(b4.Total_paralog_annotations)
b4.num_of_paralog_anno = sum(!is.na(b4.Total_paralog_annotations$ID.y))
b4.Unique_variant_annotations = b4.Total_paralog_annotations[!is.na(b4.Total_paralog_annotations$ID.y),]
b4.Unique_variant_annotations = distinct(b4.Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b4.true_num_of_paralog_anno = length(b4.Unique_variant_annotations$Variant_pos)

conf4 = conf_matrix(p4.true_num_of_paralog_anno, p3.sarco_var_align$paralog_data, b4.true_num_of_paralog_anno, b3.sarco_var_align$paralog_data)

var_removed3 = var_rem_matrix(conf3$con_table, conf4$con_table, p3.sarco_var_align$paralog_data, b3.sarco_var_align$paralog_data)

#Overall_table
Variant = c("Pathogenic variants","Benign variants","PPV","Sensitivity","Specificty","P value")
Total = c(dim(p.sarco_var_align$input_variants)[1],dim(b.sarco_var_align$input_variants)[1],"NA","NA","NA","NA")
Paralogue_Annotation_no_QC = c(format(conf$con_table[2], scientific = FALSE),format(conf$con_table[4], scientific = FALSE),round(digits = 3, conf$PPV),round(digits = 3,conf$Sensitivity),round(digits = 3,conf$Specificity),signif(digits = 3,conf$Pvalue$p.value))
Variants_remaining_after_PA_QC1 = c(format(conf2$con_table[2], scientific = FALSE),format(conf2$con_table[4], scientific = FALSE),round(digits = 3,conf2$PPV), round(digits = 3,conf2$Sensitivity),round(conf2$Specificity,digits = 3), signif(digits = 3,conf2$Pvalue$p.value))

Variants_removed_after_PA_QC1 = c(var_removed$con_table[2],var_removed$con_table[4],round(digits = 3,var_removed$PPV),round(digits = 3,var_removed$Sensitivity),round(digits = 3,var_removed$Specificity),signif(digits = 3,var_removed$Pvalue$p.value))
Variants_remaining_after_PA_QC2 = c(format(conf3$con_table[2], scientific = FALSE),format(conf3$con_table[4], scientific = FALSE),round(digits = 3,conf3$PPV),round(digits = 3,conf3$Sensitivity),round(digits = 3,conf3$Specificity),signif(digits = 3,conf3$Pvalue$p.value))
Variants_removed_after_PA_QC2 = c(sapply(c(var_removed2$con_table[2],var_removed2$con_table[4]),format,scientific=FALSE),round(digits = 3,var_removed2$PPV),round(digits = 3,var_removed2$Sensitivity),round(digits = 3,var_removed2$Specificity),signif(digits = 3,var_removed2$Pvalue$p.value))
Variants_remaining_after_PA_QC3 = c(format(conf4$con_table[2], scientific = FALSE),format(conf4$con_table[4], scientific = FALSE),round(digits = 3,conf4$PPV),round(digits = 3,conf4$Sensitivity),round(digits = 3,conf4$Specificity),signif(digits = 3,conf4$Pvalue$p.value))
Variants_removed_after_PA_QC3 = c(var_removed3$con_table[2],var_removed3$con_table[4],round(digits = 3,var_removed3$PPV),round(digits = 3,var_removed3$Sensitivity),round(digits = 3,var_removed3$Specificity),signif(digits = 3,var_removed3$Pvalue$p.value))

sarcomeric_set = data.frame(Variant,
                               Total,
                               Paralogue_Annotation_no_QC,
                               Variants_remaining_after_PA_QC1,
                               Variants_removed_after_PA_QC1,
                               Variants_remaining_after_PA_QC2,
                               Variants_removed_after_PA_QC2,
                               Variants_remaining_after_PA_QC3,
                               Variants_removed_after_PA_QC3
                               )
sarcomeric_set_t = t(sarcomeric_set)
rownames(sarcomeric_set_t)  = c("", "Total variants", "Paralogue Annotation", "Variants remaining after QC1", "Variants removed from QC1", "Variants remaining after QC2", "Variants removed after QC2", "Variants remaining after QC3", "Variants removed after QC3")
kable(sarcomeric_set_t)

```


There are however limitations to this current framework. These can be listed as the following criteria: 1)the reliance on genes with paralogues; 2) for those paralogues to have pathogenic variants; and 3) for the paralogous variants to be aligned to corresponding equivalent positions. This is not always the case. 

As a specific example, consider Hypertrophic Cardiomyopathy (HCM) and the 8 sarcomeric genes commonly associated with the genetic basis of HCM: MYH7; MYBPC3; TNNT2; TPM1; MYL2; MYL3; TNNI3; and ACTC1.

Annotating only these 8 sarcomeric genes with the whole clinvar P/LP dataset as before did not provide many annotations - without any quality control there were only `r sarcomeric_set$Paralogue_Annotation_no_QC[1]` P/LP variants and `r sarcomeric_set$Paralogue_Annotation_no_QC[2]` B/LP variants predicted to be pathogenic.

This could suggest either PA does not perform well on sarcomeric genes for the reasons stated above (paralogues to sarcomeric genes are not involed in disease) or that there is a lack of data - given more pathogenic variants to annotate with would certainly increase the likelihood of paralogous alignments. 

\newpage

```{r sarco_para, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7,fig.height=6,fig.cap="\\label{fig:sarco_para}Number of paralogues each sarcomeric gene has"}
ensembl = useMart("ENSEMBL_MART_ENSEMBL", host = "jul2018.archive.ensembl.org")
ens_human <- useDataset("hsapiens_gene_ensembl",mart=ensembl)
gene_id = c("MYH7", "MYBPC3", "TNNT2", "TPM1", "MYL2", "MYL3", "TNNI3", "ACTC1")

sarcomeric_genes = getBM(attributes = c("external_gene_name",
                                                  "hsapiens_paralog_associated_gene_name"), 
                                   filters = c("biotype", "with_hsapiens_paralog", "external_gene_name"), 
                                   values = list("protein_coding", TRUE, gene_id), mart = ens_human)

clinvar_P_LP_tableized = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized", sep = "\t", header=TRUE, stringsAsFactors=FALSE)

sarcomeric_variants = data.frame(Doubles=double(),
                                         Ints=integer(),
                                         Factors=factor(),
                                         Logicals=logical(),
                                         Characters=character(),
                                         stringsAsFactors=FALSE)
count = 1
for (variant in clinvar_P_LP_tableized$SYMBOL){
  if (variant %in% gene_id){
    sarcomeric_variants = rbind(sarcomeric_variants, clinvar_P_LP_tableized[count,])
  }
  count = count + 1
}

sarcomeric_paralog_variants = data.frame(Doubles=double(),
                                         Ints=integer(),
                                         Factors=factor(),
                                         Logicals=logical(),
                                         Characters=character(),
                                         stringsAsFactors=FALSE)
count = 1
for (variant in clinvar_P_LP_tableized$SYMBOL){
  if (variant %in% sarcomeric_genes$hsapiens_paralog_associated_gene_name){
    sarcomeric_paralog_variants = rbind(sarcomeric_paralog_variants, clinvar_P_LP_tableized[count,])
  }
  count = count + 1
}

# sarcomeric_genes_paralogs = spread(sarcomeric_genes, key = "hsapiens_paralog_associated_gene_name", "hsapiens_paralog_associated_gene_name")

sarcomeric_genes_paralogs = sarcomeric_genes %>%
  group_by(external_gene_name) %>%
  summarise(total = toString(hsapiens_paralog_associated_gene_name)) %>%
  ungroup()



kable(sarcomeric_genes_paralogs, "latex") %>%
  kable_styling(latex_options = "scale_down")
#kable(sarcomeric_paralog_variants)
```

Looking at how many paralogues the 8 sarcomeric genes have (table `sarcomeric_genes_paralogs`), there are at least 2 for each gene with MYH7 having the most - 13. This satisfies the first criteria.

In the clinvar P/LP dataset, there are `r length(sarcomeric_variants$CHROM)` known P/LP variants that lie in the 8 sarcomeric genes. But taking only the paralogues of these genes, there are only `r length(sarcomeric_paralog_variants$CHROM)` variants. Assuming the clinvar dataset is complete and considering it in isolation, this would suggest that in comparison to variants in the main 8 sarcomeric genes involved in HCM, their associated paralogous variants are not as frequently involved in disease. This still does however statisfy the second criteria.

Therefore, the third criteria is where the lack of as many annotations arises. Having additional data for both known reference variants to transfer annotations from and more query variants to transfer annotations to may resolve this as there would intuitively be more alignments available to transfer annotations. Collecting additional known pathogenic variants can be difficult as that would require established interpretation and verification of such variants. 

<!--Data in below code chunk too big. Unable to KnitR. Have to hard write numbers in paragraph.-->
```{r eval=FALSE, echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE}
#noQC
#all missense data
# p.art.paralogous_var_align = Paralogous_var_align_compressed("data/all_possible_mutations/remapped_cm_mutation_vep_canon_wRefAl.out_paraloc_paralogs2.noQC.gz", "data/all_possible_mutations/remapped_cm_mutation_vep_canon_wRefAl.out_paraloc_tableized.gz", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
gene_subset = c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1")
p.art.paralogous_var_align = Paralogous_var_align(Subset = c("Gene", gene_subset), "data/all_possible_mutations/remapped_cm_mutation_vep_canon_wRefAl.out_paraloc_paralogs2.noQC.gz", "data/all_possible_mutations/remapped_cm_mutation_vep_canon_wRefAl.out_paraloc_tableized.gz", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")



# p.art_Total_paralog_annotations = dplyr::filter(p.art.paralogous_var_align$Total_paralog_annotations, Gene %in% gene_subset)  
# p.art_num_of_paralog_anno = p.art_Total_paralog_annotations[!is.na(p.art_Total_paralog_annotations$ID.y),]
# 
# p.art_paralog_data = dplyr::filter(p.art.paralogous_var_align$paralog_data, Gene %in% gene_subset)

# patho_art_var = p.art.paralogous_var_align$Total_paralog_annotations[!is.na(p.art.paralogous_var_align$Total_paralog_annotations$ID.y),]
```

But querying every possible missense variant at all positions in the 8 sarcomeric genes can be done (30607 total variants). Doing so, predicts 1259 variants to pathogenic - a 78.7 fold increase. Though validating how many additional predictions are true positives would require additional known pathogenicity data.

```{r eval = FALSE, cache=TRUE, warning = FALSE}
##case data VEP##
case.paralogous_var_align = Paralogous_var_align("data/case_and_control/remapped_HCM_missense_LMM_OMGL_chr_wRefAl.out_paraloc_paralogs2.noQC", "data/case_and_control/remapped_HCM_missense_LMM_OMGL_chr_wRefAl.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
patho_case_var = case.paralogous_var_align$Total_paralog_annotations[!is.na(case.paralogous_var_align$Total_paralog_annotations$ID.y),]
patho_case_var_IDs = patho_case_var$ID.x
# write(patho_case_var_IDs, file = "/media/nick/Data/Users/N/Documents/PhD/Paralogues/data_files/case_controls/patho_case_var_IDs.txt", sep = "\t", ncolumns = 1)
case.data = read.csv("data/case_and_control/HCM_missense_LMM_OMGL_chr.csv", sep = ",") #only rare variants in the diseased cohort
patho_case.data = case.data[case.data$mut_id %in% patho_case_var_IDs,]
case_cases = sum(patho_case.data$sum.vs_case_count.)
Total_case_cases = sum(case.data$sum.vs_case_count.)

All_variant_case.cases = data.frame(genes = c("MYBPC3","MYH7","TNNT2","TNNI3","TPM1","ACTC1","MYL2","MYL3"), cases = c(6179,6112,6103,6047,4447,4185,4185,4185))

##control data VEP##
control.paralogous_var_align = Paralogous_var_align("data/case_and_control/remapped_ExAC_missense_sarcomeric_chr_wRefAl.out_paraloc_paralogs2.noQC", "data/case_and_control/remapped_ExAC_missense_sarcomeric_chr_wRefAl.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized") #only rare variants in the control cohort
patho_control_var = control.paralogous_var_align$Total_paralog_annotations[!is.na(control.paralogous_var_align$Total_paralog_annotations$ID.y),]
patho_control_var_IDs = patho_control_var$ID.x
# write(patho_control_var_IDs, file = "/media/nick/Data/Users/N/Documents/PhD/Paralogues/data_files/case_controls/patho_control_var_IDs.txt", sep = "\t", ncolumns = 1)
control.data = read.csv("data/case_and_control/ExAC_missense_sarcomeric_chr.csv", sep = ",")
patho_control.data = control.data[control.data$mut_id %in% patho_control_var_IDs,]
control_cases = sum(patho_control.data$mut_exac_count)
Total_control_cases = sum(control.data$mut_exac_count)

All_variant_control.cases = data.frame(genes = c("MYH7","TNNT2","TNNI3","MYBPC3","TPM1","ACTC1","MYL2","MYL3"), cases = c(60483,56800,53377,45646,58651,60271,60554,60611))

##Gene by gene analysis##
# example code for only ACTC1
# ACTC1.case.data = case.data[case.data$gene_symbol=="ACTC1",]
# patho_ACTC1.case.data = ACTC1.case.data[ACTC1.case.data$mut_id %in% patho_case_var_IDs,]
# patho_ACTC1.case.cases = sum(patho_ACTC1.case.data$sum.vs_case_count.)
# Total_ACTC1.case.cases = sum(ACTC1.case.data$sum.vs_case_count.)
#
# ACTC1.control.data = control.data[control.data$gene_symbol=="ACTC1",]
# patho_ACTC1.control.data = ACTC1.control.data[ACTC1.control.data$mut_id %in% patho_control_var_IDs,]
# patho_ACTC1.control.cases = sum(patho_ACTC1.control.data$mut_exac_count)
# Total_ACTC1.control.cases = sum(ACTC1.control.data$mut_exac_count)
# calc_EF(Total_ACTC1.case.cases,All_variant_case.cases[All_variant_case.cases$genes=="ACTC1","cases"],Total_ACTC1.control.cases,All_variant_control.cases[All_variant_control.cases$genes=="ACTC1","cases"]) #all rare variants
# calc_EF(patho_ACTC1.case.cases,All_variant_case.cases[All_variant_case.cases$genes=="ACTC1","cases"],patho_ACTC1.control.cases,All_variant_control.cases[All_variant_control.cases$genes=="ACTC1","cases"]) #rare variants predicted to be pathogenic by PA 

# x = case_control_gene_split(case.data, patho_case_var_IDs)
# y = case_control_gene_split(control.data, patho_control_var_IDs, "ACTC1")

# gene_EFs = list()

gene_EFs = data.frame()
for (gene in c("MYH7","MYBPC3","TNNT2","TPM1","MYL2","MYL3","TNNI3","ACTC1")){
# for (gene in c("MYBPC3")){
  cases = case_control_gene_split(case.data, patho_case_var_IDs, gene)
  controls = case_control_gene_split(control.data, patho_control_var_IDs, gene)
  all_rare_vars = calc_EF(cases$Total_gene_cases,
                          All_variant_case.cases[All_variant_case.cases$genes==gene,"cases"]-cases$Total_gene_cases,
                          controls$Total_gene_cases,
                          All_variant_control.cases[All_variant_control.cases$genes==gene,"cases"]-controls$Total_gene_cases)
  pa_predicted_rare_vars = calc_EF(cases$pa_predicted_gene_cases,
                                   All_variant_case.cases[All_variant_case.cases$genes==gene,"cases"]-cases$pa_predicted_gene_cases,
                                   controls$pa_predicted_gene_cases,
                                   All_variant_control.cases[All_variant_control.cases$genes==gene,"cases"]-controls$pa_predicted_gene_cases)
  # gene_EFs[[gene]] = list("all_rare_vars" = all_rare_vars, "pa_predicted_rare_vars" = pa_predicted_rare_vars)
  # tmp_df = data.frame(gene,toString(all_rare_vars$OR),toString(all_rare_vars$OR_CI),toString(all_rare_vars$EF),toString(all_rare_vars$EF_CI),toString(pa_predicted_rare_vars$OR),toString(pa_predicted_rare_vars$OR_CI),toString(pa_predicted_rare_vars$EF),toString(pa_predicted_rare_vars$EF_CI))
  tmp_df = data.frame(gene,cell_format(all_rare_vars$OR,all_rare_vars$OR_CI),cell_format(all_rare_vars$EF,all_rare_vars$EF_CI),cell_format(pa_predicted_rare_vars$OR,pa_predicted_rare_vars$OR_CI),cell_format(pa_predicted_rare_vars$EF,pa_predicted_rare_vars$EF_CI))
  gene_EFs = dplyr::bind_rows(gene_EFs,tmp_df)
}
##Overall##
all_rare_vars = calc_EF(Total_case_cases,
                        max(All_variant_case.cases$cases)-Total_case_cases,
                        Total_control_cases,
                        max(All_variant_control.cases$cases)-Total_control_cases)
pa_predicted_rare_vars = calc_EF(case_cases,
                                 max(All_variant_case.cases$cases)-case_cases,
                                 control_cases,
                                 max(All_variant_control.cases$cases)-control_cases)
gene_EFs = dplyr::bind_rows(gene_EFs,data.frame(gene="Overall",
                                                cell_format(all_rare_vars$OR,all_rare_vars$OR_CI),
                                                cell_format(all_rare_vars$EF,all_rare_vars$EF_CI),
                                                cell_format(pa_predicted_rare_vars$OR,pa_predicted_rare_vars$OR_CI),
                                                cell_format(pa_predicted_rare_vars$EF,pa_predicted_rare_vars$EF_CI)
                                                )
                            )
# gene_EFs = dplyr::bind_rows(gene_EFs,data.frame(gene="Overall",
#                                                 # sapply(c(
#                                                 toString(all_rare_vars$OR),
#                                                 toString(all_rare_vars$OR_CI),
#                                                 toString(all_rare_vars$EF),
#                                                 toString(all_rare_vars$EF_CI),
#                                                 toString(pa_predicted_rare_vars$OR),
#                                                 toString(pa_predicted_rare_vars$OR_CI),
#                                                 toString(pa_predicted_rare_vars$EF),
#                                                 toString(pa_predicted_rare_vars$EF_CI)
#                                                 # ),format,nsmall=3)
#                                                 )
#                             )
colnames(gene_EFs) = c("Gene","All rare variants OR","All rare variants EF","Predicted pathogenic OR","Predicted pathogenic EF")
write.table(gene_EFs, file = "data/case_and_control/gene_EFs.csv", na="", row.names = FALSE, col.names = TRUE, sep = ",", quote=FALSE)

```

However, it must be noted that even in genes with few predictions, paralogue annotation can still be functional and such predictions applicable. In the case of the 8 sarcomeric genes, one can still validate the use of paralogue annotation in a case control study.

Cases with positive (bad) outcome
Number in exposed group: 	
39
 	
Number in control group: 	
28
 	
Cases with negative (good) outcome
Number in exposed group: 	
41404
 	
Number in control group: 	
456365

Results
Odds ratio 	15.3524
95 % CI:	9.4468 to 24.9499
z statistic	11.024
Significance level	P < 0.0001

EFs show that even with few predictions, PA still works.

For example in the 8 sarcomeric genes involved in HCM [MYH7, MYBPC3, TNNT2, TPM1, MYL2, MYL3, TNNI3, ACTC1], taking MYH7 there were not many paralogous alignments.   since most paralogues of HCM disease genes are no.

Infact performing the analysis on all possible missesnse mutations for these set of genes still shows a lack of annotation...

Hence we calculated EFs in order to see for those few variants that are predicted to be pathogenic, how often do they appear to be causative of disease in a disease cohort case control study. Segway to HCM validation.

\newpage
## Calculations of EFs for LQTS
There are 15 genes that are associated with LQTS but the main 3 are: KCNH2; KCNQ1; and SCN5A.

The other 12 are: AKAP9; ANK2; CACNA1C; CALM1; CALM2;	CAV3;	KCNE1; KCNE2; KCNJ2; KCNJ5; SCN4B; and SNTA1. [@alders2018long]

For the OMGL data, Total disease cases would be considered as all variants in the spreadsheet. Need to first filter out non missense variants. Then take rare variants (Rare variants defined as https://www.nature.com/articles/gim201726; disease specific - in this case its LQTS; each disease has different threshold of "rarity") and variants that are predicted pathogenic by PA. Look at the same genes in gnomad, take variants from only exome (ask roddy - basically case data can only be extracted from vcfs https://gnomad.broadinstitute.org/downloads or from older version of gnomad browser). Extract only rare variants and ones predicted pathogenic by PA. Calculate EFs from this.

The Total case counts for gnomad would be the mean of all allele numbers of respective variants being considered divided by 2 (due to homo/heterozygous). I.e. for each variant take the AN number from gnomad vcf and sum them together, take the mean and divide by 2.

Using https://www.nature.com/articles/gim201726/tables/1 and using the upper CI level, e.g. for LQTS = 8.2 x 106. If the maximum filtering allele frequency (max out of all the sub populations, i.e. non-global) is smaller than 8.2 x 10-6 then that variant is considered "rare"

OMGL data
```{r eval=FALSE, echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE}
#Each line is a single patient count with the variant. exacAllelecount is the cell that has the same variant exac count
#For any variants where exac allele count is NA, just put as 0? Check this assumption

KCNQ1_data_ref = read.csv("data/LQTS/KCNQ1_JamesWare_270117.csv", sep = ",")
colnames(KCNQ1_data_ref)[1] = "gene"
KCNQ1_data = KCNQ1_data_ref[KCNQ1_data_ref$codingEffect == "missense",]
KCNQ1_gNomen_count = as.data.frame(table(KCNQ1_data$gNomen))
KCNQ1_gNomen_count = KCNQ1_gNomen_count[KCNQ1_gNomen_count$Freq > 0,]
colnames(KCNQ1_gNomen_count)[2] = "sum.vs_case_count"
KCNQ1_data = left_join(KCNQ1_data, KCNQ1_gNomen_count, by = c("gNomen" = "Var1"))
KCNQ1_data$exacAlleleCount[is.na(KCNQ1_data$exacAlleleCount)] = 0
unique_KCNQ1_gNomen = KCNQ1_data[match(unique(KCNQ1_data$gNomen), KCNQ1_data$gNomen),]

KCNH2_data_ref = read.csv("data/LQTS/KCNH2_JamesWare_270117.csv", sep = ",") 
# KCNH2_data = KCNH2_data[,c("gene", "chrom", "codingEffect", "assembly", "gDNAstart", "gDNAend", "pNomen", "pathogenicityClass", "rsId", "exacQuality", "exacFilter", "exacAlleleCount", "exacAllFreq", "exacAFRFreq", "exacAMRFreq", "exacEASFreq", "exacSASFreq", "exacNFEFreq", "exacFINFreq", "exacOTHFreq", "exacDP")]
KCNH2_data_ref = KCNH2_data_ref[,!(names(KCNH2_data_ref) %in% setdiff(colnames(KCNH2_data_ref), colnames(KCNQ1_data_ref)))]
KCNH2_data = KCNH2_data_ref[KCNH2_data_ref$codingEffect == "missense",]
KCNH2_gNomen_count = as.data.frame(table(KCNH2_data$gNomen))
KCNH2_gNomen_count = KCNH2_gNomen_count[KCNH2_gNomen_count$Freq > 0,]
colnames(KCNH2_gNomen_count)[2] = "sum.vs_case_count"
KCNH2_data = left_join(KCNH2_data, KCNH2_gNomen_count, by = c("gNomen" = "Var1"))
KCNH2_data$exacAlleleCount[is.na(KCNH2_data$exacAlleleCount)] = 0
unique_KCNH2_gNomen = KCNH2_data[match(unique(KCNH2_data$gNomen), KCNH2_data$gNomen),]

SCN5A_data_ref = read.csv("data/LQTS/SCN5A_JamesWare_270117.csv", sep = ",")
colnames(SCN5A_data_ref)[1] = "gene"
SCN5A_data_ref = SCN5A_data_ref[,!(names(SCN5A_data_ref) %in% setdiff(colnames(SCN5A_data_ref), colnames(KCNQ1_data_ref)))]
SCN5A_data = SCN5A_data_ref[SCN5A_data_ref$codingEffect == "missense",]
SCN5A_gNomen_count = as.data.frame(table(SCN5A_data$gNomen))
SCN5A_gNomen_count = SCN5A_gNomen_count[SCN5A_gNomen_count$Freq > 0,]
colnames(SCN5A_gNomen_count)[2] = "sum.vs_case_count"
SCN5A_data = left_join(SCN5A_data, SCN5A_gNomen_count, by = c("gNomen" = "Var1"))
SCN5A_data$exacAlleleCount[is.na(SCN5A_data$exacAlleleCount)] = 0
unique_SCN5A_gNomen = SCN5A_data[match(unique(SCN5A_data$gNomen), SCN5A_data$gNomen),]


Total_var = do.call("rbind", list(unique_KCNH2_gNomen, unique_KCNQ1_gNomen, unique_SCN5A_gNomen))
vcf_Total_var = Total_var[,c("chrom", "gDNAstart", "gNomen")]
vcf_Total_var$gNomen = as.character(vcf_Total_var$gNomen)
vcf_Total_var$REF = sapply(strsplit(vcf_Total_var$gNomen, ">"), "[", 1)
vcf_Total_var$REF = substr(vcf_Total_var$REF, nchar(vcf_Total_var$REF), nchar(vcf_Total_var$REF))
vcf_Total_var$ALT = sapply(strsplit(vcf_Total_var$gNomen, ">"), "[", 2)
vcf_Total_var$QUAl = "."
vcf_Total_var$FILTER = "."
vcf_Total_var$INFO = "."
colnames(vcf_Total_var)[1:3] = c("#CHROM", "POS", "ID")
write.table(vcf_Total_var, file = "data/LQTS/All_variants.vcf", sep = "\t", quote = FALSE, row.names = FALSE)
```

Ackerman data from ICC_Mutations
```{r eval=FALSE, echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE}

#Finding only rare variants
disease_cohort_variants = read.csv("data/LQTS/Mayo_LQTS_variants_for_vcf_convertion.vcf", sep = "\t", stringsAsFactors = 0)

gnomad_total_variants = read.csv("data/LQTS/gnomAD_v2.1.1_LQTS_variants.vcf", sep = "\t", stringsAsFactors = 0)
gnomad_rare_variants = read.csv("data/LQTS/gnomAD_v2.1.1_LQTS_variants_rare.vcf", sep = "\t", stringsAsFactors = 0) #RUN PA ON THIS FILE

rare_disease_cohort_variants = data.frame()
# disease_cohort_variants = disease_cohort_variants[1:10,]
for (row in 1:nrow(disease_cohort_variants)){ #FINDING ONLY RARE VARIANTS BASED ON GNOMAD FREQUENCY
  if (
    paste0(disease_cohort_variants[row,"X.CHROM"], disease_cohort_variants[row,"POS"], disease_cohort_variants[row,"REF"], disease_cohort_variants[row,"ALT"]) %in% paste0(gnomad_rare_variants$X.CHROM, gnomad_rare_variants$POS, gnomad_rare_variants$REF, gnomad_rare_variants$ALT)
    ){
    rare_disease_cohort_variants = rbind(rare_disease_cohort_variants, disease_cohort_variants[row,])
  } else if (!(
    paste0(disease_cohort_variants[row,"X.CHROM"], disease_cohort_variants[row,"POS"], disease_cohort_variants[row,"REF"], disease_cohort_variants[row,"ALT"]) %in% paste0(gnomad_total_variants$X.CHROM, gnomad_total_variants$POS, gnomad_total_variants$REF, gnomad_total_variants$ALT)
  )){
    rare_disease_cohort_variants = rbind(rare_disease_cohort_variants, disease_cohort_variants[row,])
  }
}
colnames(rare_disease_cohort_variants)[1] = "#CHROM"
write.table(rare_disease_cohort_variants, file = "data/LQTS/rare_disease_cohort_variants.vcf", na="", row.names = FALSE, col.names = TRUE, sep = "\t", quote=FALSE) #RUN PA ON THIS FILE


###disease cohort
disease_data_ref = read.csv("data/LQTS/Mayo_LQTS_variants.csv", sep = ",", stringsAsFactors = 0)
disease_data_ref$cases = sapply(disease_data_ref$rep_patients,strsplit,"/")
disease_data_ref$cases = sapply(disease_data_ref$cases,function(x) x[1])
disease_data_ref$total_cases = sapply(disease_data_ref$rep_patients,strsplit,"/")
disease_data_ref$total_cases = sapply(disease_data_ref$total_cases,function(x) x[2])
#total cases sequenced for each gene
disease_total_variant_cases = data.frame(genes = c("KCNE1","KCNE2","KCNH2","KCNQ1","SCN5A"), cases = c(2500,2500,2500,2500,2500))
#individual cases for rare variants
KCNE1_disease_rare_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNE1" & disease_data_ref$mut_id %in% rare_disease_cohort_variants$ID, "cases"]))
KCNE2_disease_rare_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNE2" & disease_data_ref$mut_id %in% rare_disease_cohort_variants$ID, "cases"]))
KCNH2_disease_rare_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNH2" & disease_data_ref$mut_id %in% rare_disease_cohort_variants$ID, "cases"]))
KCNQ1_disease_rare_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNQ1" & disease_data_ref$mut_id %in% rare_disease_cohort_variants$ID, "cases"]))
SCN5A_disease_rare_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "SCN5A" & disease_data_ref$mut_id %in% rare_disease_cohort_variants$ID, "cases"]))
#individual cases for variants predicted pathogenic by PA
disease.paralogous_var_align = Paralogous_var_align("data/LQTS/rare_disease_cohort_variants.out_paraloc_paralogs2.noQC", "data/LQTS/rare_disease_cohort_variants.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
disease_cohort_data = disease_data_ref[disease_data_ref$mut_id %in% disease.paralogous_var_align$Unique_variant_annotations$ID.x,]
KCNE1_disease_pa_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNE1" & disease_data_ref$mut_id %in% disease.paralogous_var_align$Unique_variant_annotations$ID.x, "cases"]))
KCNE2_disease_pa_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNE2" & disease_data_ref$mut_id %in% disease.paralogous_var_align$Unique_variant_annotations$ID.x, "cases"]))
KCNH2_disease_pa_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNH2" & disease_data_ref$mut_id %in% disease.paralogous_var_align$Unique_variant_annotations$ID.x, "cases"]))
KCNQ1_disease_pa_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "KCNQ1" & disease_data_ref$mut_id %in% disease.paralogous_var_align$Unique_variant_annotations$ID.x, "cases"]))
SCN5A_disease_pa_variant_cases = sum(as.numeric(disease_data_ref[disease_data_ref$gene_symbol == "SCN5A" & disease_data_ref$mut_id %in% disease.paralogous_var_align$Unique_variant_annotations$ID.x, "cases"]))

###control cohort
control_data_ref = read.csv("data/LQTS/gnomAD_v2.1.1_LQTS_variants_with_customIDs.csv", sep = ",", stringsAsFactors = 0)
control_data_ref  = control_data_ref[control_data_ref$Gene != "Gene",] #remove the extra header lines from when concatenated the separate gene csv files together

gnomad_total_variants$X.CHROM = as.character(gnomad_total_variants$X.CHROM)
gnomad_total_variants$POS = as.character(gnomad_total_variants$POS)
control_data_ref = full_join(control_data_ref,gnomad_total_variants,by = c("Chromosome" = "X.CHROM", "Position" = "POS", "Reference" = "REF", "Alternate" = "ALT")) #IN ORDER TO FIX THE GODDAM CUSTOM IDS, JOIN WITH ORIGINAL CSV FILE AND USE ID INSTEAD OF rsID

KCNE1_control_total_variant_cases = control_data_ref[control_data_ref$Gene == "KCNE1",]
KCNE2_control_total_variant_cases = control_data_ref[control_data_ref$Gene == "KCNE2",]
KCNH2_control_total_variant_cases = control_data_ref[control_data_ref$Gene == "KCNH2",]
KCNQ1_control_total_variant_cases = control_data_ref[control_data_ref$Gene == "KCNQ1",]
SCN5A_control_total_variant_cases = control_data_ref[control_data_ref$Gene == "SCN5A",]
#total cases sequence for each gene
control_total_variant_cases = data.frame(
  genes = c("KCNE1","KCNE2","KCNH2","KCNQ1","SCN5A"), 
  cases = c(
    mean(as.numeric(KCNE1_control_total_variant_cases$Allele.Number))/2,
    mean(as.numeric(KCNE2_control_total_variant_cases$Allele.Number))/2,
    mean(as.numeric(KCNH2_control_total_variant_cases$Allele.Number))/2,
    mean(as.numeric(KCNQ1_control_total_variant_cases$Allele.Number))/2,
    mean(as.numeric(SCN5A_control_total_variant_cases$Allele.Number))/2))
#individual cases for rare variants
KCNE1_control_rare_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNE1" & control_data_ref$ID %in% gnomad_rare_variants$ID, "Allele.Count"]))
KCNE2_control_rare_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNE2" & control_data_ref$ID %in% gnomad_rare_variants$ID, "Allele.Count"]))
KCNH2_control_rare_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNH2" & control_data_ref$ID %in% gnomad_rare_variants$ID, "Allele.Count"]))
KCNQ1_control_rare_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNQ1" & control_data_ref$ID %in% gnomad_rare_variants$ID, "Allele.Count"]))
SCN5A_control_rare_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "SCN5A" & control_data_ref$ID %in% gnomad_rare_variants$ID, "Allele.Count"]))
#individual cases for variants predicted pathogenic by PA
control.paralogous_var_align = Paralogous_var_align("data/LQTS/gnomAD_v2.1.1_LQTS_variants_rare.out_paraloc_paralogs2.noQC", "data/LQTS/gnomAD_v2.1.1_LQTS_variants_rare.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized") 
control_cohort_data = control_data_ref[control_data_ref$ID %in% control.paralogous_var_align$Unique_variant_annotations$ID.x,] 
KCNE1_control_pa_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNE1" & control_data_ref$ID %in% control.paralogous_var_align$Unique_variant_annotations$ID.x, "Allele.Count"]))
KCNE2_control_pa_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNE2" & control_data_ref$ID %in% control.paralogous_var_align$Unique_variant_annotations$ID.x, "Allele.Count"]))
KCNH2_control_pa_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNH2" & control_data_ref$ID %in% control.paralogous_var_align$Unique_variant_annotations$ID.x, "Allele.Count"]))
KCNQ1_control_pa_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "KCNQ1" & control_data_ref$ID %in% control.paralogous_var_align$Unique_variant_annotations$ID.x, "Allele.Count"]))
SCN5A_control_pa_variant_cases = sum(as.numeric(control_data_ref[control_data_ref$Gene == "SCN5A" & control_data_ref$ID %in% control.paralogous_var_align$Unique_variant_annotations$ID.x, "Allele.Count"]))


###Calculating EFs
KCNE1_rare_EF = calc_EF(
  KCNE1_disease_rare_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNE1", "cases"]-KCNE1_disease_rare_variant_cases,
  KCNE1_control_rare_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNE1", "cases"]-KCNE1_control_rare_variant_cases
  )
KCNE1_pa_EF = calc_EF(
  KCNE1_disease_pa_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNE1", "cases"]-KCNE1_disease_pa_variant_cases,
  KCNE1_control_pa_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNE1", "cases"]-KCNE1_control_pa_variant_cases
  )

KCNE2_rare_EF = calc_EF(
  KCNE2_disease_rare_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNE2", "cases"]-KCNE2_disease_rare_variant_cases,
  KCNE2_control_rare_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNE2", "cases"]-KCNE2_control_rare_variant_cases
  )
KCNE2_pa_EF = calc_EF(
  KCNE2_disease_pa_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNE2", "cases"]-KCNE2_disease_pa_variant_cases,
  KCNE2_control_pa_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNE2", "cases"]-KCNE2_control_pa_variant_cases
  )

KCNH2_rare_EF = calc_EF(
  KCNH2_disease_rare_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNH2", "cases"]-KCNH2_disease_rare_variant_cases,
  KCNH2_control_rare_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNH2", "cases"]-KCNH2_control_rare_variant_cases
  )
KCNH2_pa_EF = calc_EF(
  KCNH2_disease_pa_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNH2", "cases"]-KCNH2_disease_pa_variant_cases,
  KCNH2_control_pa_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNH2", "cases"]-KCNH2_control_pa_variant_cases
  )

KCNQ1_rare_EF = calc_EF(
  KCNQ1_disease_rare_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNQ1", "cases"]-KCNQ1_disease_rare_variant_cases,
  KCNQ1_control_rare_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNQ1", "cases"]-KCNQ1_control_rare_variant_cases
  )
KCNQ1_pa_EF = calc_EF(
  KCNQ1_disease_pa_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "KCNQ1", "cases"]-KCNQ1_disease_pa_variant_cases,
  KCNQ1_control_pa_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "KCNQ1", "cases"]-KCNQ1_control_pa_variant_cases
  )

SCN5A_rare_EF = calc_EF(
  SCN5A_disease_rare_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "SCN5A", "cases"]-SCN5A_disease_rare_variant_cases,
  SCN5A_control_rare_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "SCN5A", "cases"]-SCN5A_control_rare_variant_cases
  )
SCN5A_pa_EF = calc_EF(
  SCN5A_disease_pa_variant_cases,
  disease_total_variant_cases[disease_total_variant_cases$genes == "SCN5A", "cases"]-SCN5A_disease_pa_variant_cases,
  SCN5A_control_pa_variant_cases,
  control_total_variant_cases[control_total_variant_cases$genes == "SCN5A", "cases"]-SCN5A_control_pa_variant_cases
  )

gene_EFs = data.frame(
  gene = c("KCNE1","KCNE2","KCNH2","KCNQ1","SCN5A"),
  rare_variant_OR = c(cell_format(KCNE1_rare_EF$OR, KCNE1_rare_EF$OR_CI),
                      cell_format(KCNE2_rare_EF$OR, KCNE2_rare_EF$OR_CI),
                      cell_format(KCNH2_rare_EF$OR, KCNH2_rare_EF$OR_CI),
                      cell_format(KCNQ1_rare_EF$OR, KCNQ1_rare_EF$OR_CI),
                      cell_format(SCN5A_rare_EF$OR, SCN5A_rare_EF$OR_CI)),
  rare_variant_EF = c(cell_format(KCNE1_rare_EF$EF, KCNE1_rare_EF$EF_CI),
                      cell_format(KCNE2_rare_EF$EF, KCNE2_rare_EF$EF_CI),
                      cell_format(KCNH2_rare_EF$EF, KCNH2_rare_EF$EF_CI),
                      cell_format(KCNQ1_rare_EF$EF, KCNQ1_rare_EF$EF_CI),
                      cell_format(SCN5A_rare_EF$EF, SCN5A_rare_EF$EF_CI)),
  pa_variant_OR = c(cell_format(KCNE1_pa_EF$OR, KCNE1_pa_EF$OR_CI),
                    cell_format(KCNE2_pa_EF$OR, KCNE2_pa_EF$OR_CI),
                    cell_format(KCNH2_pa_EF$OR, KCNH2_pa_EF$OR_CI),
                    cell_format(KCNQ1_pa_EF$OR, KCNQ1_pa_EF$OR_CI),
                    cell_format(SCN5A_pa_EF$OR, SCN5A_pa_EF$OR_CI)),
  pa_variant_EF = c(cell_format(KCNE1_pa_EF$EF, KCNE1_pa_EF$EF_CI),
                    cell_format(KCNE2_pa_EF$EF, KCNE2_pa_EF$EF_CI),
                    cell_format(KCNH2_pa_EF$EF, KCNH2_pa_EF$EF_CI),
                    cell_format(KCNQ1_pa_EF$EF, KCNQ1_pa_EF$EF_CI),
                    cell_format(SCN5A_pa_EF$EF, SCN5A_pa_EF$EF_CI))
  )
write.table(gene_EFs, file = "data/LQTS/Ackerman_LQTS_EFs.csv", na="", row.names = FALSE, col.names = FALSE, sep = ",", quote=FALSE)


# disease_cohort_variants$check_in_rare_gnomad = ifelse(
#   is.na(
#     match(
#       paste0(disease_cohort_variants$X.CHROM, disease_cohort_variants$POS, disease_cohort_variants$REF, disease_cohort_variants$ALT),
#       paste0(gnomad_rare_variants$X.CHROM, gnomad_rare_variants$POS, gnomad_rare_variants$REF, gnomad_rare_variants$ALT)
#       )
#     ),"No", "Yes"
#   )
# 
# disease_cohort_variants$check_in_total_gnomad

# disease_cohort_variants_tmp = disease_cohort_variants[,c(1,2,4,5)]
# gnomad_rare_variants_tmp = gnomad_rare_variants[,c(1,2,4,5)]

# vcf_conv = read.csv("data/LQTS/Mayo_LQTS_variants_for_vcf_convertion.csv", sep = ",", stringsAsFactors = 0)
```


## Comparison to other tools
Looking at clinvar variants:
```{r cf_sift_revel, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE}
# #First initial run to see what variants are not in overlapping genes
# # p.normal_PA2 = Paralogous_var_align_no_overlap("data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# p.normal_initial = Paralogous_var_align(Overlap = 0, "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# # b.normal_PA2 = Paralogous_var_align_no_overlap("data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_initial = Paralogous_var_align(Overlap = 0, "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_initial_conf = conf_matrix(p.normal_initial$true_num_of_paralog_anno, p.normal_initial$input_variants, b.normal_initial$true_num_of_paralog_anno, b.normal_initial$input_variants)
# 
# 
# ##Preparing and annotating with other classifiers
# #revel - as revel has the least number of variants annotated, all other methods were made sure to use the same variants that revel managed to annotate
# Revel_tableized_file = read.csv(file="data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized_w_REVEL", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
# Revel_tableized_data = Revel_tableized_file
# Revel_tableized_data$Variant_pos = paste(Revel_tableized_data$CHROM,Revel_tableized_data$POS, sep = " ")
# Revel_tableized_data = Revel_tableized_data[!is.na(Revel_tableized_data$Codons)&!is.na(Revel_tableized_data$Paralogue_Vars),]
# Revel_tableized_data = filter(Revel_tableized_data, ID %in% p.normal_initial$n_occur$Var1)
# p.num_of_revel = length(which(Revel_tableized_data$REVEL_score > 0.5))
# 
# b.Revel_tableized_file = read.csv(file="data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized_w_REVEL", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
# b.Revel_tableized_data = b.Revel_tableized_file
# b.Revel_tableized_data$Variant_pos = paste(b.Revel_tableized_data$CHROM,b.Revel_tableized_data$POS, sep = " ")
# b.Revel_tableized_data = b.Revel_tableized_data[!is.na(b.Revel_tableized_data$Codons)&!is.na(b.Revel_tableized_data$Paralogue_Vars),]
# b.Revel_tableized_data = filter(b.Revel_tableized_data, ID %in% b.normal_initial$n_occur$Var1)
# b.num_of_revel = length(which(b.Revel_tableized_data$REVEL_score > 0.5))
# 
# # Revel_TP = p.num_of_revel
# # Revel_FP = b.num_of_revel
# # Revel_TN = nrow(b.Revel_tableized_data)-b.num_of_revel
# # Revel_FN = nrow(Revel_tableized_data)-p.num_of_revel
# # Revel_PPV = Revel_TP/(Revel_TP+Revel_FP)
# # Revel_Sensitivty = Revel_TP/(Revel_TP+Revel_FN)
# # Revel_Specificity = Revel_TN/(Revel_TN+Revel_FP)
# # Revel_FPR = Revel_FP/(Revel_FP+Revel_TN)
# # Revel_ACC = (Revel_TP + Revel_TN)/(Revel_TP + Revel_TN + Revel_FP + Revel_FN)
# # Revel_con_table = matrix(
# #   c(nrow(Revel_tableized_data),
# #     p.num_of_revel,
# #     nrow(b.Revel_tableized_data),
# #     b.num_of_revel
# #     ), ncol = 2
# # )
# # colnames(Revel_con_table) = c("Pathogenic", "Benign")
# # rownames(Revel_con_table) = c("Number of variants in total", "Number of variants predicted as pathogenic")
# 
# revel_conf = raw_conf_matrix(p.num_of_revel,Revel_tableized_data,b.num_of_revel,b.Revel_tableized_data)
# 
# #sift
# Sift_tableized_file = read.csv(file="data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc.hg19_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
# Sift_tableized_data = Sift_tableized_file
# Sift_tableized_data$Variant_pos = paste(Sift_tableized_data$CHROM,Sift_tableized_data$POS, sep = " ")
# 
# Sift_tableized_data = filter(Sift_tableized_data, ID %in% Revel_tableized_data$ID)
# p.num_of_sift = length(which(Sift_tableized_data$SIFT_pred == "D")) 
# 
# b.Sift_tableized_file = read.csv(file="data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc.hg19_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
# b.Sift_tableized_data = b.Sift_tableized_file
# b.Sift_tableized_data$Variant_pos = paste(b.Sift_tableized_data$CHROM,b.Sift_tableized_data$POS, sep = " ")
# 
# b.Sift_tableized_data = filter(b.Sift_tableized_data, ID %in% b.Revel_tableized_data$ID)
# b.num_of_sift = length(which(b.Sift_tableized_data$SIFT_pred == "D")) 
# 
# # Sift_TP = p.num_of_sift
# # Sift_FP = b.num_of_sift
# # Sift_TN = nrow(b.Sift_tableized_data)-b.num_of_sift
# # Sift_FN = nrow(Sift_tableized_data)-p.num_of_sift
# # Sift_PPV = Sift_TP/(Sift_TP+Sift_FP)
# # Sift_Sensitivty = Sift_TP/(Sift_TP+Sift_FN)
# # Sift_Specificity = Sift_TN/(Sift_TN+Sift_FP)
# # Sift_FPR = Sift_FP/(Sift_FP+Sift_TN)
# # Sift_ACC = (Sift_TP + Sift_TN)/(Sift_TP + Sift_TN + Sift_FP + Sift_FN)
# # Sift_con_table = matrix(
# #   c(nrow(Sift_tableized_data),
# #     p.num_of_sift,
# #     nrow(b.Sift_tableized_data),
# #     b.num_of_sift
# #     ), ncol = 2
# # )
# # colnames(Sift_con_table) = c("Pathogenic", "Benign")
# # rownames(Sift_con_table) = c("Number of variants in total", "Number of variants predicted as pathogenic")
# 
# sift_conf = raw_conf_matrix(p.num_of_sift,Sift_tableized_data,b.num_of_sift,b.Sift_tableized_data)
# 
# #M-CAP
# Mcap_tableized_file = read.csv(file="data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc.hg19_multianno.vcf_tableized_org_M_CAP", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
# Mcap_tableized_data = filter(Mcap_tableized_file, ID %in% Revel_tableized_data$ID)
# p.num_of_mcap = length(which(Mcap_tableized_data$M.CAP_pred == "D"))
# 
# b.Mcap_tableized_file = read.csv(file="data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc.hg19_multianno.vcf_tableized_org_M_CAP", sep = "\t", header=TRUE, stringsAsFactors=FALSE)
# b.Mcap_tableized_data = filter(b.Mcap_tableized_file, ID %in% b.Revel_tableized_data$ID)
# b.num_of_mcap = length(which(b.Mcap_tableized_data$M.CAP_pred == "D")) 
#  
# # Mcap_TP = p.num_of_mcap
# # Mcap_FP = b.num_of_mcap
# # Mcap_TN = nrow(b.Mcap_tableized_data)-b.num_of_mcap
# # Mcap_FN = nrow(Mcap_tableized_data)-p.num_of_mcap
# # Mcap_PPV = Mcap_TP/(Mcap_TP+Mcap_FP)
# # Mcap_Sensitivty = Mcap_TP/(Mcap_TP+Mcap_FN)
# # Mcap_Specificity = Mcap_TN/(Mcap_TN+Mcap_FP)
# # Mcap_FPR = Mcap_FP/(Mcap_FP+Mcap_TN)
# # Mcap_ACC = (Mcap_TP + Mcap_TN)/(Mcap_TP + Mcap_TN + Mcap_FP + Mcap_FN)
# # Mcap_con_table = matrix(
# #   c(nrow(Mcap_tableized_data),
# #     p.num_of_mcap,
# #     nrow(b.Mcap_tableized_data),
# #     b.num_of_mcap
# #     ), ncol = 2
# # )
# # colnames(Mcap_con_table) = c("Pathogenic", "Benign")
# # rownames(Mcap_con_table) = c("Number of variants in total", "Number of variants predicted as pathogenic")
# 
# mcap_conf = raw_conf_matrix(p.num_of_mcap,Mcap_tableized_data,b.num_of_mcap,b.Mcap_tableized_data)


##New build 38 data
p.normal_initial = Paralogous_var_align(Overlap = 0, "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")

other_tools_file = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs_no_empty_lines.hg38_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
other_tools_file = other_tools_file[complete.cases(other_tools_file),]
other_tools_file = filter(other_tools_file, ID %in% p.normal_initial$n_occur$Var1)
p.num_of_revel = length(which(other_tools_file$REVEL_score > 0.5))
p.num_of_sift = length(which(other_tools_file$SIFT_pred == "D"))
p.num_of_mcap = length(which(other_tools_file$M.CAP_pred == "D"))
p.num_of_cadd = length(which(other_tools_file$CADD_phred > 15))

b.normal_initial = Paralogous_var_align(Overlap = 0, "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")

b.other_tools_file = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs_no_empty_lines.hg38_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
b.other_tools_file = b.other_tools_file[complete.cases(b.other_tools_file),]
b.other_tools_file = filter(b.other_tools_file, ID %in% b.normal_initial$n_occur$Var1)
b.num_of_revel = length(which(b.other_tools_file$REVEL_score > 0.5))
b.num_of_sift = length(which(b.other_tools_file$SIFT_pred == "D"))
b.num_of_mcap = length(which(b.other_tools_file$M.CAP_pred == "D"))
b.num_of_cadd = length(which(b.other_tools_file$CADD_phred > 15))

sift_conf = raw_conf_matrix(p.num_of_sift,other_tools_file,b.num_of_sift,b.other_tools_file)
mcap_conf = raw_conf_matrix(p.num_of_mcap,other_tools_file,b.num_of_mcap,b.other_tools_file)
revel_conf = raw_conf_matrix(p.num_of_revel,other_tools_file,b.num_of_revel,b.other_tools_file)
cadd_conf = raw_conf_matrix(p.num_of_cadd,other_tools_file,b.num_of_cadd,b.other_tools_file)

sift1_conf = conf_matrix(p.num_of_sift,other_tools_file,b.num_of_sift,b.other_tools_file)
mcap1_conf = conf_matrix(p.num_of_mcap,other_tools_file,b.num_of_mcap,b.other_tools_file)
revel1_conf = conf_matrix(p.num_of_revel,other_tools_file,b.num_of_revel,b.other_tools_file)
cadd1_conf = conf_matrix(p.num_of_cadd,other_tools_file,b.num_of_cadd,b.other_tools_file)


#PA
# p.normal_PA = Paralogous_var_align(Subset = c("ID",Revel_tableized_data$ID),"data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA = Paralogous_var_align(Subset = c("ID",b.Revel_tableized_data$ID),"data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA_conf = conf_matrix(p.normal_PA$true_num_of_paralog_anno, p.normal_PA$input_variants, b.normal_PA$true_num_of_paralog_anno, b.normal_PA$input_variants)

p.normal_PA = Paralogous_var_align(Subset = c("ID",other_tools_file$ID),"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA = Paralogous_var_align(Subset = c("ID",b.other_tools_file$ID),"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
normal_PA_conf = conf_matrix(p.normal_PA$true_num_of_paralog_anno, p.normal_PA$input_variants, b.normal_PA$true_num_of_paralog_anno, b.normal_PA$input_variants)




# p.normal_PA2 = Paralogous_var_align(Subset = c("ID",Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA2 = Paralogous_var_align(Subset = c("ID",b.Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA2_conf = conf_matrix(p.normal_PA2$true_num_of_paralog_anno, p.normal_PA2$input_variants, b.normal_PA2$true_num_of_paralog_anno, b.normal_PA2$input_variants)

p.normal_PA2 = Paralogous_var_align(Subset = c("ID",other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA2 = Paralogous_var_align(Subset = c("ID",b.other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
normal_PA2_conf = conf_matrix(p.normal_PA2$true_num_of_paralog_anno, p.normal_PA2$input_variants, b.normal_PA2$true_num_of_paralog_anno, b.normal_PA2$input_variants)


# p.normal_PA2 = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_paralogs2.para_con", "data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA2 = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_paralogs2.para_con", "data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA2_conf = conf_matrix(p.normal_PA2$true_num_of_paralog_anno, p.normal_PA2$input_variants, b.normal_PA2$true_num_of_paralog_anno, b.normal_PA2$input_variants)

# p.normal_PA3 = Paralogous_var_align(Subset = c("ID",Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA3 = Paralogous_var_align(Subset = c("ID",b.Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA3_conf = conf_matrix(p.normal_PA3$true_num_of_paralog_anno, p.normal_PA3$input_variants, b.normal_PA3$true_num_of_paralog_anno, b.normal_PA3$input_variants)

p.normal_PA3 = Paralogous_var_align(Subset = c("ID",other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA3 = Paralogous_var_align(Subset = c("ID",b.other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
normal_PA3_conf = conf_matrix(p.normal_PA3$true_num_of_paralog_anno, p.normal_PA3$input_variants, b.normal_PA3$true_num_of_paralog_anno, b.normal_PA3$input_variants)


p.normal_PA4 = list()
p.normal_PA4$Total_paralog_annotations = p.normal_PA3$Total_paralog_annotations[p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
p.normal_PA4$Total_paralog_annotations = distinct(p.normal_PA4$Total_paralog_annotations)
p.normal_PA4$Unique_variant_annotations = p.normal_PA4$Total_paralog_annotations[!is.na(p.normal_PA4$Total_paralog_annotations$ID.y),]
p.normal_PA4$Unique_variant_annotations = distinct(p.normal_PA4$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p.normal_PA4$true_num_of_paralog_anno = length(p.normal_PA4$Unique_variant_annotations$Variant_pos)
b.normal_PA4 = list()
b.normal_PA4$Total_paralog_annotations = b.normal_PA3$Total_paralog_annotations[b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA4$Total_paralog_annotations = distinct(b.normal_PA4$Total_paralog_annotations)
b.normal_PA4$Unique_variant_annotations = b.normal_PA4$Total_paralog_annotations[!is.na(b.normal_PA4$Total_paralog_annotations$ID.y),]
b.normal_PA4$Unique_variant_annotations = distinct(b.normal_PA4$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b.normal_PA4$true_num_of_paralog_anno = length(b.normal_PA4$Unique_variant_annotations$Variant_pos)
normal_PA4_conf = conf_matrix(p.normal_PA4$true_num_of_paralog_anno, p.normal_PA3$input_variants, b.normal_PA4$true_num_of_paralog_anno, b.normal_PA3$input_variants)

#QC3+QC1, no QC2  
p.normal_PA5 = list()
p.normal_PA5$Total_paralog_annotations = p.normal_PA2$Total_paralog_annotations[p.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.y,]
p.normal_PA5$Total_paralog_annotations = distinct(p.normal_PA5$Total_paralog_annotations)
p.normal_PA5$Unique_variant_annotations = p.normal_PA5$Total_paralog_annotations[!is.na(p.normal_PA5$Total_paralog_annotations$ID.y),]
p.normal_PA5$Unique_variant_annotations = distinct(p.normal_PA5$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p.normal_PA5$true_num_of_paralog_anno = length(p.normal_PA5$Unique_variant_annotations$Variant_pos)
b.normal_PA5 = list()
b.normal_PA5$Total_paralog_annotations = b.normal_PA2$Total_paralog_annotations[b.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA5$Total_paralog_annotations = distinct(b.normal_PA5$Total_paralog_annotations)
b.normal_PA5$Unique_variant_annotations = b.normal_PA5$Total_paralog_annotations[!is.na(b.normal_PA5$Total_paralog_annotations$ID.y),]
b.normal_PA5$Unique_variant_annotations = distinct(b.normal_PA5$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b.normal_PA5$true_num_of_paralog_anno = length(b.normal_PA5$Unique_variant_annotations$Variant_pos)
normal_PA5_conf = conf_matrix(p.normal_PA5$true_num_of_paralog_anno, p.normal_PA2$input_variants, b.normal_PA5$true_num_of_paralog_anno, b.normal_PA2$input_variants)

# p.normal_PA = Subset_var_align_id(Revel_tableized_data$ID,"data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA = Subset_var_align_id(b.Revel_tableized_data$ID,"data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA_conf = conf_matrix(p.normal_PA$num_of_paralog_anno, p.normal_PA$paralog_data, b.normal_PA$num_of_paralog_anno, b.normal_PA$paralog_data)
# p.normal_PA2 = Subset_var_align_id(Revel_tableized_data$ID,"data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA2 = Subset_var_align_id(b.Revel_tableized_data$ID,"data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA2_conf = conf_matrix(p.normal_PA2$num_of_paralog_anno, p.normal_PA2$paralog_data, b.normal_PA2$num_of_paralog_anno, b.normal_PA2$paralog_data)
# p.normal_PA3 = Subset_var_align_id(Revel_tableized_data$ID,"data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# b.normal_PA3 = Subset_var_align_id(b.Revel_tableized_data$ID,"data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
# normal_PA3_conf = conf_matrix(p.normal_PA3$num_of_paralog_anno, p.normal_PA3$paralog_data, b.normal_PA3$num_of_paralog_anno, b.normal_PA3$paralog_data)

#ParaZ
Sensitivity_list = vector()
Sens_CI_list = vector()
FPR_list = vector()
con_table_list = matrix(, ncol = 2)
PPV_list = vector()
PPV_CI_list = vector()
P_value_list = vector()
Specificity_list = vector()
##Table##
paraz_set = t(data.frame(Total = c(revel_conf$con_table[1],revel_conf$con_table[3],"NA","NA","NA","NA","NA","NA")))
colnames(paraz_set) = c("Pathogenic variants","Benign variants","PPV","PPV CI","Sensitivity","Sensitivity CI","Specificity","P value")
####
for (i in seq(0,11)){
#noQC
#pathogenic to pathogenic
# p.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
  # p.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
  p.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",other_tools_file$ID),
                                           # "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC",
                                           "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
  
#benign to pathogenic
# b.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
  # b.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",b.Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
  b.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",b.other_tools_file$ID),
                                           # "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC",
                                           "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
  
  
  paraz.conf = conf_matrix(p.paraz_var_align$true_num_of_paralog_anno, p.paraz_var_align$paralog_data, b.paraz_var_align$true_num_of_paralog_anno, b.paraz_var_align$paralog_data)
  Sensitivity_list = c(Sensitivity_list, paraz.conf$Sensitivity)
  Sens_CI_list = c(Sens_CI_list, 1.96*paraz.conf$Sensitivity_SE)
  FPR_list = c(FPR_list, paraz.conf$FPR)
  con_table_list = rbind(con_table_list, paraz.conf$con_table)
  PPV_list = c(PPV_list, paraz.conf$PPV)
  PPV_CI_list= c(PPV_CI_list, 1.96*paraz.conf$PPV_SE)
  P_value_list = c(P_value_list, paraz.conf$Pvalue$p.value)
  Specificity_list = c(Specificity_list, paraz.conf$Specificity)
  
  PA_threshold_row = c(format(paraz.conf$con_table[2], scientific = FALSE), format(paraz.conf$con_table[2,2], scientific = FALSE), round(paraz.conf$PPV, digits = 3), round(1.96*paraz.conf$PPV_SE, digits = 3), round(paraz.conf$Sensitivity, digits = 3), round(1.96*paraz.conf$Sensitivity_SE, digits =3), round(paraz.conf$Specificity,digits = 3), signif(paraz.conf$Pvalue$p.value, digits = 3))
  paraz_set = rbind(paraz_set, PA_threshold_row)
}

# Overall dataframe
other_tools_set = t(data.frame(
  # Total = c(sapply(c(revel_conf$con_table[1],revel_conf$con_table[3]),format,scientific = FALSE),"NA","NA","NA","NA"),
  Total = c(sapply(c(dim(other_tools_file)[1],dim(b.other_tools_file)[1]),format,scientific = FALSE),"NA","NA","NA","NA","NA","NA"),
  Sift = c(sapply(c(sift_conf$con_table[2],sift_conf$con_table[4]),format,scientific = FALSE),
           sapply(c(sift_conf$PPV,sift1_conf$PPV_SE*1.96,sift_conf$Sensitivity,sift1_conf$Sensitivity_SE*1.96,sift_conf$Specificity,sift_conf$Pvalue$p.value),signif,digits=3)),
  MCap = c(sapply(c(mcap_conf$con_table[2],mcap_conf$con_table[4]),format,scientific = FALSE),
           sapply(c(mcap_conf$PPV,mcap1_conf$PPV_SE*1.96,mcap_conf$Sensitivity,mcap1_conf$Sensitivity_SE*1.96,mcap_conf$Specificity,mcap_conf$Pvalue$p.value),signif,digits=3)),
  Cadd = c(sapply(c(cadd_conf$con_table[2],cadd_conf$con_table[4]),format,scientific = FALSE),
           sapply(c(cadd_conf$PPV,cadd1_conf$PPV_SE*1.96,cadd_conf$Sensitivity,cadd1_conf$Sensitivity_SE*1.96,cadd_conf$Specificity,cadd_conf$Pvalue$p.value),signif,digits=3)),
  Revel = c(sapply(c(revel_conf$con_table[2],revel_conf$con_table[4]),format,scientific = FALSE),
            sapply(c(revel_conf$PPV,revel1_conf$PPV_SE*1.96,revel_conf$Sensitivity,revel1_conf$Sensitivity_SE*1.96,revel_conf$Specificity,revel_conf$Pvalue$p.value),signif,digits=3)),
  PA.noQC = c(sapply(c(normal_PA_conf$con_table[2],normal_PA_conf$con_table[4]),format,scientific = FALSE),
         sapply(c(normal_PA_conf$PPV,normal_PA_conf$PPV_SE*1.96,normal_PA_conf$Sensitivity,normal_PA_conf$Sensitivity_SE*1.96,normal_PA_conf$Specificity,normal_PA_conf$Pvalue$p.value),signif,digits=3)),
  PA.para_con = c(sapply(c(normal_PA2_conf$con_table[2],normal_PA2_conf$con_table[4]),format,scientific = FALSE),
                  sapply(c(normal_PA2_conf$PPV,normal_PA2_conf$PPV_SE*1.96,normal_PA2_conf$Sensitivity,normal_PA2_conf$Sensitivity_SE*1.96,normal_PA2_conf$Specificity,normal_PA2_conf$Pvalue$p.value),signif,digits=3)),
  PA.all_con = c(sapply(c(normal_PA3_conf$con_table[2],normal_PA3_conf$con_table[4]),format,scientific = FALSE),
                 sapply(c(normal_PA3_conf$PPV,normal_PA3_conf$PPV_SE*1.96,normal_PA3_conf$Sensitivity,normal_PA3_conf$Sensitivity_SE*1.96,normal_PA3_conf$Specificity,normal_PA3_conf$Pvalue$p.value),signif,digits=3)),
  PA.alt_con = c(sapply(c(normal_PA4_conf$con_table[2],normal_PA4_conf$con_table[4]),format,scientific = FALSE),
                 sapply(c(normal_PA4_conf$PPV,normal_PA4_conf$PPV_SE*1.96,normal_PA4_conf$Sensitivity,normal_PA4_conf$Sensitivity_SE*1.96,normal_PA4_conf$Specificity,normal_PA4_conf$Pvalue$p.value),signif,digits=3)),
  PA.alt_con_w_para_con_no_all_con = c(sapply(c(normal_PA5_conf$con_table[2],normal_PA5_conf$con_table[4]),format,scientific = FALSE),
                 sapply(c(normal_PA5_conf$PPV,normal_PA5_conf$PPV_SE*1.96,normal_PA5_conf$Sensitivity,normal_PA5_conf$Sensitivity_SE*1.96,normal_PA5_conf$Specificity,normal_PA5_conf$Pvalue$p.value),signif,digits=3))
))
colnames(other_tools_set) = c("Pathogenic variants","Benign variants","PPV","PPV CI","Sensitivity","Sens CI","Specificity","P value")
paraz_set = paraz_set[-1,]
other_tools_set = rbind(other_tools_set,paraz_set)
rownames(other_tools_set) = c("Total","Sift","MCap","CADD","Revel","Paralogue Annotation","Paralogue Annotation (QC1)","Paralogue Annotation (QC2)", "Paralogue Annotation (QC3)", "Paralogue Annotation (QC3, no QC2)", "Para z_score cutoff 0","Para z_score cutoff 1","Para z_score cutoff 2","Para z_score cutoff 3","Para z_score cutoff 4","Para z_score cutoff 5","Para z_score cutoff 6","Para z_score cutoff 7","Para z_score cutoff 8","Para z_score cutoff 9","Para z_score cutoff 10","Para z_score cutoff 11")
# write.table(other_tools_set, file = "data/other_tools_set.csv", na="", row.names = TRUE, col.names = TRUE, sep = ",", quote=FALSE)
write.table(other_tools_set, file = "data/other_tools_set_w_paraz_afterQC1.csv", na="", row.names = TRUE, col.names = TRUE, sep = ",", quote=FALSE)


#for subsetting the massive dataframe above for easier graph making
other_tools_data = data.frame(
  Precision = c(as.numeric(other_tools_set[2:9,3])),
  Precision_CI = c(as.numeric(other_tools_set[2:9,4])),
  Sensitivity = c(as.numeric(other_tools_set[2:9,5])),
  Sensitivity_CI = c(as.numeric(other_tools_set[2:9,6]))
  )
other_tools_data$Type = c("Sift", "MCap", "CADD", "Revel", "noQC", "QC1" ,"QC2", "QC3")
# other_tools_data = rbind(other_tools_data, prc_data2[1:4,])
other_tools_data$Type = factor(other_tools_data$Type, levels = c("Sift", "MCap", "CADD", "Revel", "noQC", "QC1" ,"QC2", "QC3"))

gg4 = ggplot(other_tools_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=0) +
  scale_shape_manual(values=c(0, 1, 2, 5, 15, 16, 17, 18)) +
  xlim(0,1) + ylim(0.8,1)  + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation\ncompared with other tools", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5)) #+ scale_color_discrete(breaks=c("Sift","MCap","Revel","noQC","QC1","QC2","QC3"))
ggsave("images/PRC_with_other_tools_plot.png")

other_tools_data = data.frame(
  Precision = c(as.numeric(other_tools_set[2:10,3])),
  Precision_CI = c(as.numeric(other_tools_set[2:10,4])),
  Sensitivity = c(as.numeric(other_tools_set[2:10,5])),
  Sensitivity_CI = c(as.numeric(other_tools_set[2:10,6]))
  )
other_tools_data$Type = c("Sift", "MCap", "CADD", "Revel", "noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)")
# other_tools_data = rbind(other_tools_data, prc_data2[1:4,])
other_tools_data$Type = factor(other_tools_data$Type, levels = c("Sift", "MCap", "CADD", "Revel", "noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)"))

gg5 = ggplot(other_tools_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=0) +
  scale_shape_manual(values=c(0, 1, 2, 5, 15, 16, 17, 18,19)) +
  xlim(0,1) + ylim(0.8,1)  + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation\ncompared with other tools", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5)) #+ scale_color_discrete(breaks=c("Sift","MCap","Revel","noQC","QC1","QC2","QC3"))
ggsave("images/PRC_with_other_tools_with_QC3_QC1_no_QC2.png")

other_tools_data = data.frame(
  Precision = c(as.numeric(other_tools_set[6:10,3])),
  Precision_CI = c(as.numeric(other_tools_set[6:10,4])),
  Sensitivity = c(as.numeric(other_tools_set[6:10,5])),
  Sensitivity_CI = c(as.numeric(other_tools_set[6:10,6]))
  )
other_tools_data$Type = c("noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)")
# other_tools_data = rbind(other_tools_data, prc_data2[1:4,])
other_tools_data$Type = factor(other_tools_data$Type, levels = c("noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)"))

gg6 = ggplot(other_tools_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=0) +
  scale_shape_manual(values=c(15, 16, 17, 18, 19)) +
  xlim(0,1) + ylim(0.8,1)  + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation\ncompared with other tools", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5)) #+ scale_color_discrete(breaks=c("Sift","MCap","Revel","noQC","QC1","QC2","QC3"))
ggsave("images/PRC_with_QC3_QC1_no_QC2.png")

```

Looking at only SCN5A:
```{r cf_sift_revel_SCN%A, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE}
##New build 38 data
p.normal_initial = Paralogous_var_align(Subset = c("Gene", "SCN5A"),Overlap = 0, "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")

other_tools_file = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs_no_empty_lines.hg38_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
other_tools_file = other_tools_file[complete.cases(other_tools_file),]
other_tools_file = filter(other_tools_file, ID %in% p.normal_initial$n_occur$Var1)
p.num_of_revel = length(which(other_tools_file$REVEL_score > 0.5))
p.num_of_sift = length(which(other_tools_file$SIFT_pred == "D"))
p.num_of_mcap = length(which(other_tools_file$M.CAP_pred == "D"))
p.num_of_cadd = length(which(other_tools_file$CADD_phred > 15))

b.normal_initial = Paralogous_var_align(Subset = c("Gene", "SCN5A"),Overlap = 0, "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized")

b.other_tools_file = read.csv(file="data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs_no_empty_lines.hg38_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
b.other_tools_file = b.other_tools_file[complete.cases(b.other_tools_file),]
b.other_tools_file = filter(b.other_tools_file, ID %in% b.normal_initial$n_occur$Var1)
b.num_of_revel = length(which(b.other_tools_file$REVEL_score > 0.5))
b.num_of_sift = length(which(b.other_tools_file$SIFT_pred == "D"))
b.num_of_mcap = length(which(b.other_tools_file$M.CAP_pred == "D"))
b.num_of_cadd = length(which(b.other_tools_file$CADD_phred > 15))

sift1_conf = conf_matrix(p.num_of_sift,other_tools_file,b.num_of_sift,b.other_tools_file)
mcap1_conf = conf_matrix(p.num_of_mcap,other_tools_file,b.num_of_mcap,b.other_tools_file)
revel1_conf = conf_matrix(p.num_of_revel,other_tools_file,b.num_of_revel,b.other_tools_file)
cadd1_conf = conf_matrix(p.num_of_cadd,other_tools_file,b.num_of_cadd,b.other_tools_file)


#PA


p.normal_PA = Paralogous_var_align(Subset = c("ID",other_tools_file$ID),"data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA = Paralogous_var_align(Subset = c("ID",b.other_tools_file$ID),"data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
normal_PA_conf = conf_matrix(p.normal_PA$true_num_of_paralog_anno, p.normal_PA$input_variants, b.normal_PA$true_num_of_paralog_anno, b.normal_PA$input_variants)



p.normal_PA2 = Paralogous_var_align(Subset = c("ID",other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA2 = Paralogous_var_align(Subset = c("ID",b.other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.para_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
normal_PA2_conf = conf_matrix(p.normal_PA2$true_num_of_paralog_anno, p.normal_PA2$input_variants, b.normal_PA2$true_num_of_paralog_anno, b.normal_PA2$input_variants)


p.normal_PA3 = Paralogous_var_align(Subset = c("ID",other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA3 = Paralogous_var_align(Subset = c("ID",b.other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.all_con", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
normal_PA3_conf = conf_matrix(p.normal_PA3$true_num_of_paralog_anno, p.normal_PA3$input_variants, b.normal_PA3$true_num_of_paralog_anno, b.normal_PA3$input_variants)


p.normal_PA4 = list()
p.normal_PA4$Total_paralog_annotations = p.normal_PA3$Total_paralog_annotations[p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
p.normal_PA4$Total_paralog_annotations = distinct(p.normal_PA4$Total_paralog_annotations)
p.normal_PA4$Unique_variant_annotations = p.normal_PA4$Total_paralog_annotations[!is.na(p.normal_PA4$Total_paralog_annotations$ID.y),]
p.normal_PA4$Unique_variant_annotations = distinct(p.normal_PA4$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p.normal_PA4$true_num_of_paralog_anno = length(p.normal_PA4$Unique_variant_annotations$Variant_pos)
b.normal_PA4 = list()
b.normal_PA4$Total_paralog_annotations = b.normal_PA3$Total_paralog_annotations[b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA4$Total_paralog_annotations = distinct(b.normal_PA4$Total_paralog_annotations)
b.normal_PA4$Unique_variant_annotations = b.normal_PA4$Total_paralog_annotations[!is.na(b.normal_PA4$Total_paralog_annotations$ID.y),]
b.normal_PA4$Unique_variant_annotations = distinct(b.normal_PA4$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b.normal_PA4$true_num_of_paralog_anno = length(b.normal_PA4$Unique_variant_annotations$Variant_pos)
normal_PA4_conf = conf_matrix(p.normal_PA4$true_num_of_paralog_anno, p.normal_PA3$input_variants, b.normal_PA4$true_num_of_paralog_anno, b.normal_PA3$input_variants)

p.normal_PA5 = list()
p.normal_PA5$Total_paralog_annotations = p.normal_PA2$Total_paralog_annotations[p.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.y,]
p.normal_PA5$Total_paralog_annotations = distinct(p.normal_PA5$Total_paralog_annotations)
p.normal_PA5$Unique_variant_annotations = p.normal_PA5$Total_paralog_annotations[!is.na(p.normal_PA5$Total_paralog_annotations$ID.y),]
p.normal_PA5$Unique_variant_annotations = distinct(p.normal_PA5$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
p.normal_PA5$true_num_of_paralog_anno = length(p.normal_PA5$Unique_variant_annotations$Variant_pos)
b.normal_PA5 = list()
b.normal_PA5$Total_paralog_annotations = b.normal_PA2$Total_paralog_annotations[b.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA5$Total_paralog_annotations = distinct(b.normal_PA5$Total_paralog_annotations)
b.normal_PA5$Unique_variant_annotations = b.normal_PA5$Total_paralog_annotations[!is.na(b.normal_PA5$Total_paralog_annotations$ID.y),]
b.normal_PA5$Unique_variant_annotations = distinct(b.normal_PA5$Unique_variant_annotations[c("Variant_pos", "ID.x", "REF.x", "ALT.x")])
b.normal_PA5$true_num_of_paralog_anno = length(b.normal_PA5$Unique_variant_annotations$Variant_pos)
normal_PA5_conf = conf_matrix(p.normal_PA5$true_num_of_paralog_anno, p.normal_PA2$input_variants, b.normal_PA5$true_num_of_paralog_anno, b.normal_PA2$input_variants)

# #ParaZ
# Sensitivity_list = vector()
# FPR_list = vector()
# con_table_list = matrix(, ncol = 2)
# PPV_list = vector()
# P_value_list = vector()
# Specificity_list = vector()
# ##Table##
# paraz_set = t(data.frame(Total = c(revel_conf$con_table[1],revel_conf$con_table[3],"NA","NA","NA","NA")))
# colnames(paraz_set) = c("Pathogenic variants","Benign variants","PPV","Sensitivity","Specificity","P value")
# ####
# for (i in seq(0,11)){
# #noQC
# #pathogenic to pathogenic
# # p.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
#   # p.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
#   p.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
#   
# #benign to pathogenic
# # b.paraz_var_align = ParaZ_var_align(i,"data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_alleles.single.b38.Benign_and_LikelyBenign.out_paraloc_tableized", "data/clinvar/clinvar_alleles.single.b38.Pathogenic_and_LikelyPathogenic.out_paraloc_tableized")
#   # b.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",b.Revel_tableized_data$ID), "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20181028_GRCh37_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
#   b.paraz_var_align = Paralogous_var_align(paraz_cutoff = i, Subset = c("ID",b.other_tools_file$ID), "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_paralogs2.noQC", "data/clinvar/clinvar_20190114_GRCh38_onlyBenign_and_Likely_benign.out_paraloc_tableized", "data/clinvar/clinvar_20190114_GRCh38_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
#   
#   
#   paraz.conf = conf_matrix(p.paraz_var_align$true_num_of_paralog_anno, p.paraz_var_align$paralog_data, b.paraz_var_align$true_num_of_paralog_anno, b.paraz_var_align$paralog_data)
#   Sensitivity_list = c(Sensitivity_list, paraz.conf$Sensitivity)
#   FPR_list = c(FPR_list, paraz.conf$FPR)
#   con_table_list = rbind(con_table_list, paraz.conf$con_table)
#   PPV_list = c(PPV_list, paraz.conf$PPV)
#   P_value_list = c(P_value_list, paraz.conf$Pvalue$p.value)
#   Specificity_list = c(Specificity_list, paraz.conf$Specificity)
#   
#   PA_threshold_row = c(format(paraz.conf$con_table[2], scientific = FALSE), format(paraz.conf$con_table[2,2], scientific = FALSE), round(paraz.conf$PPV, digits = 3), round(paraz.conf$Sensitivity, digits = 3), round(paraz.conf$Specificity,digits = 3), signif(paraz.conf$Pvalue$p.value, digits = 3))
#   paraz_set = rbind(paraz_set, PA_threshold_row)
# }

# Overall dataframe
other_tools_set = t(data.frame(
  # Total = c(sapply(c(revel_conf$con_table[1],revel_conf$con_table[3]),format,scientific = FALSE),"NA","NA","NA","NA"),
  Total = c(sapply(c(dim(other_tools_file)[1],dim(b.other_tools_file)[1]),format,scientific = FALSE),"NA","NA","NA","NA","NA","NA"),
  Sift = c(sapply(c(sift1_conf$con_table[2],sift1_conf$con_table[4]),format,scientific = FALSE),
           sapply(c(sift1_conf$PPV,sift1_conf$PPV_SE*1.96,sift1_conf$Sensitivity,sift1_conf$Sensitivity_SE*1.96,sift1_conf$Specificity,sift1_conf$Pvalue$p.value),signif,digits=3)),
  MCap = c(sapply(c(mcap1_conf$con_table[2],mcap1_conf$con_table[4]),format,scientific = FALSE),
           sapply(c(mcap1_conf$PPV,mcap1_conf$PPV_SE*1.96,mcap1_conf$Sensitivity,mcap1_conf$Sensitivity_SE*1.96,mcap1_conf$Specificity,mcap1_conf$Pvalue$p.value),signif,digits=3)),
  Cadd = c(sapply(c(cadd1_conf$con_table[2],cadd1_conf$con_table[4]),format,scientific = FALSE),
           sapply(c(cadd1_conf$PPV,cadd1_conf$PPV_SE*1.96,cadd1_conf$Sensitivity,cadd1_conf$Sensitivity_SE*1.96,cadd1_conf$Specificity,cadd1_conf$Pvalue$p.value),signif,digits=3)),
  Revel = c(sapply(c(revel1_conf$con_table[2],revel1_conf$con_table[4]),format,scientific = FALSE),
            sapply(c(revel1_conf$PPV,revel1_conf$PPV_SE*1.96,revel1_conf$Sensitivity,revel1_conf$Sensitivity_SE*1.96,revel1_conf$Specificity,revel1_conf$Pvalue$p.value),signif,digits=3)),
  PA.noQC = c(sapply(c(normal_PA_conf$con_table[2],normal_PA_conf$con_table[4]),format,scientific = FALSE),
         sapply(c(normal_PA_conf$PPV,normal_PA_conf$PPV_SE*1.96,normal_PA_conf$Sensitivity,normal_PA_conf$Sensitivity_SE*1.96,normal_PA_conf$Specificity,normal_PA_conf$Pvalue$p.value),signif,digits=3)),
  PA.para_con = c(sapply(c(normal_PA2_conf$con_table[2],normal_PA2_conf$con_table[4]),format,scientific = FALSE),
                  sapply(c(normal_PA2_conf$PPV,normal_PA2_conf$PPV_SE*1.96,normal_PA2_conf$Sensitivity,normal_PA2_conf$Sensitivity_SE*1.96,normal_PA2_conf$Specificity,normal_PA2_conf$Pvalue$p.value),signif,digits=3)),
  PA.all_con = c(sapply(c(normal_PA3_conf$con_table[2],normal_PA3_conf$con_table[4]),format,scientific = FALSE),
                 sapply(c(normal_PA3_conf$PPV,normal_PA3_conf$PPV_SE*1.96,normal_PA3_conf$Sensitivity,normal_PA3_conf$Sensitivity_SE*1.96,normal_PA3_conf$Specificity,normal_PA3_conf$Pvalue$p.value),signif,digits=3)),
  PA.alt_con = c(sapply(c(normal_PA4_conf$con_table[2],normal_PA4_conf$con_table[4]),format,scientific = FALSE),
                 sapply(c(normal_PA4_conf$PPV,normal_PA4_conf$PPV_SE*1.96,normal_PA4_conf$Sensitivity,normal_PA4_conf$Sensitivity_SE*1.96,normal_PA4_conf$Specificity,normal_PA4_conf$Pvalue$p.value),signif,digits=3)),
  PA.alt_con_w_QC1_no_QC2 = c(sapply(c(normal_PA5_conf$con_table[2],normal_PA5_conf$con_table[4]),format,scientific = FALSE),
                 sapply(c(normal_PA5_conf$PPV,normal_PA5_conf$PPV_SE*1.96,normal_PA5_conf$Sensitivity,normal_PA5_conf$Sensitivity_SE*1.96,normal_PA5_conf$Specificity,normal_PA5_conf$Pvalue$p.value),signif,digits=3))
))
colnames(other_tools_set) = c("Pathogenic variants","Benign variants","PPV","PPV CI","Sensitivity","Sens CI","Specificity","P value")
##REMINDER TO FIX APPENDING OF PARAZ SET AS HAVE NOT YET ADDED CI FOR PPV AND SENS
# paraz_set = paraz_set[-1,]
# other_tools_set = rbind(other_tools_set,paraz_set)
# rownames(other_tools_set) = c("Total","Sift","MCap","CADD","Revel","Paralogue Annotation","Paralogue Annotation (QC1)","Paralogue Annotation (QC2)", "Paralogue Annotation (QC3)","Para z_score cutoff 0","Para z_score cutoff 1","Para z_score cutoff 2","Para z_score cutoff 3","Para z_score cutoff 4","Para z_score cutoff 5","Para z_score cutoff 6","Para z_score cutoff 7","Para z_score cutoff 8","Para z_score cutoff 9","Para z_score cutoff 10","Para z_score cutoff 11")
# write.table(other_tools_set, file = "data/other_tools_set.csv", na="", row.names = TRUE, col.names = TRUE, sep = ",", quote=FALSE)
# 

other_tools_data = data.frame(
  Precision = c(as.numeric(other_tools_set[6:10,3])),
  Precision_CI = c(as.numeric(other_tools_set[6:10,4])),
  Sensitivity = c(as.numeric(other_tools_set[6:10,5])),
  Sensitivity_CI = c(as.numeric(other_tools_set[6:10,6]))
  )
other_tools_data$Type = c("noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)")
# other_tools_data = rbind(other_tools_data, prc_data2[1:4,])
other_tools_data$Type = factor(other_tools_data$Type, levels = c("noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)"))

gg = ggplot(other_tools_data, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=0) +
  scale_shape_manual(values=c(15, 16, 17, 18, 19)) +
  xlim(0,1) + ylim(0.75,1)  + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5)) 
ggsave("images/PRC_of_SCN5A.png")

```


Looking at UniFun and TP53-TA variants:
```{r cf_UniFun_TP53-TA, cache=TRUE, echo = FALSE, results='asis', message = FALSE, warning=FALSE}
p.normal_PA = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_paralogs2.noQC", "data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_paralogs2.noQC", "data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
p.normal_PA2 = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_paralogs2.para_con", "data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA2 = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_paralogs2.para_con", "data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
p.normal_PA3 = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_paralogs2.all_con", "data/TP53-TA/4704961/TP53-TA_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA3 = Paralogous_var_align("data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_paralogs2.all_con", "data/TP53-TA/4704961/TP53-TA_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
p.normal_PA4 = list()
p.normal_PA4$Total_paralog_annotations = p.normal_PA3$Total_paralog_annotations[p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA4 = list()
b.normal_PA4$Total_paralog_annotations = b.normal_PA3$Total_paralog_annotations[b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
normal_PA4 = Alt_con_PA(p.normal_PA4$Total_paralog_annotations,b.normal_PA4$Total_paralog_annotations)
p.normal_PA5 = list()
p.normal_PA5$Total_paralog_annotations = p.normal_PA2$Total_paralog_annotations[p.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA5 = list()
b.normal_PA5$Total_paralog_annotations = b.normal_PA2$Total_paralog_annotations[b.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA2$Total_paralog_annotations$ALT_Amino_acids.y,]
normal_PA5 = Alt_con_PA(p.normal_PA5$Total_paralog_annotations,b.normal_PA5$Total_paralog_annotations)

other_tools = read.csv(file="data/TP53-TA/4704961/TP53-TA_deleterious.vcf.hg19_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
p.num_of_revel = length(which(other_tools$REVEL_score > 0.5))
p.num_of_sift = length(which(other_tools$SIFT_pred == "D"))
p.num_of_mcap = length(which(other_tools$M.CAP_pred == "D"))
p.num_of_cadd = length(which(other_tools$CADD_phred > 15))

b.other_tools = read.csv(file="data/TP53-TA/4704961/TP53-TA_benign.vcf.hg19_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
b.num_of_revel = length(which(b.other_tools$REVEL_score > 0.5))
b.num_of_sift = length(which(b.other_tools$SIFT_pred == "D"))
b.num_of_mcap = length(which(b.other_tools$M.CAP_pred == "D"))
b.num_of_cadd = length(which(b.other_tools$CADD_phred > 15))


# sift_conf = raw_conf_matrix(p.num_of_sift,other_tools,b.num_of_sift,b.other_tools)
# mcap_conf = raw_conf_matrix(p.num_of_mcap,other_tools,b.num_of_mcap,b.other_tools)
# revel_conf = raw_conf_matrix(p.num_of_revel,other_tools,b.num_of_revel,b.other_tools)
# cadd_conf = raw_conf_matrix(p.num_of_cadd,other_tools,b.num_of_cadd,b.other_tools)
sift1_conf = conf_matrix(p.num_of_sift,other_tools,b.num_of_sift,b.other_tools)
mcap1_conf = conf_matrix(p.num_of_mcap,other_tools,b.num_of_mcap,b.other_tools)
revel1_conf = conf_matrix(p.num_of_revel,other_tools,b.num_of_revel,b.other_tools)
cadd1_conf = conf_matrix(p.num_of_cadd,other_tools,b.num_of_cadd,b.other_tools)

normal_PA_conf = conf_matrix(p.normal_PA$true_num_of_paralog_anno, other_tools, b.normal_PA$true_num_of_paralog_anno, b.other_tools)
normal_PA2_conf = conf_matrix(p.normal_PA2$true_num_of_paralog_anno, other_tools, b.normal_PA2$true_num_of_paralog_anno, b.other_tools)
normal_PA3_conf = conf_matrix(p.normal_PA3$true_num_of_paralog_anno, other_tools, b.normal_PA3$true_num_of_paralog_anno, b.other_tools)
normal_PA4_conf = conf_matrix(normal_PA4$p.normal_PA4$true_num_of_paralog_anno, other_tools, normal_PA4$b.normal_PA4$true_num_of_paralog_anno, b.other_tools)
normal_PA5_conf = conf_matrix(normal_PA5$p.normal_PA4$true_num_of_paralog_anno, other_tools, normal_PA5$b.normal_PA4$true_num_of_paralog_anno, b.other_tools)

df1 = data.frame(
  Type = c("Sift","Mcap","Cadd","Revel","noQC","QC1","QC2","QC3", "QC3 (skipped QC2)"),
  Precision = c(sift1_conf$PPV,mcap1_conf$PPV,cadd1_conf$PPV,revel1_conf$PPV,normal_PA_conf$PPV,normal_PA2_conf$PPV,normal_PA3_conf$PPV,normal_PA4_conf$PPV,normal_PA5_conf$PPV),
  Precision_CI = c(sift1_conf$PPV_SE*.196,mcap1_conf$PPV_SE*1.96,cadd1_conf$PPV_SE*1.96,revel1_conf$PPV_SE*1.96,normal_PA_conf$PPV_SE*1.96,normal_PA2_conf$PPV_SE*1.96,normal_PA3_conf$PPV_SE*1.96,normal_PA4_conf$PPV_SE*1.96,normal_PA5_conf$PPV_SE*1.96),
  Sensitivity = c(sift1_conf$Sensitivity,mcap1_conf$Sensitivity,cadd1_conf$Sensitivity,revel1_conf$Sensitivity,normal_PA_conf$Sensitivity,normal_PA2_conf$Sensitivity,normal_PA3_conf$Sensitivity,normal_PA4_conf$Sensitivity,normal_PA5_conf$Sensitivity),
  Sensitivity_CI = c(sift1_conf$Sensitivity_SE*1.96,mcap1_conf$Sensitivity_SE*1.96,cadd1_conf$Sensitivity_SE*1.96,revel1_conf$Sensitivity_SE*1.96,normal_PA_conf$Sensitivity_SE*1.96,normal_PA2_conf$Sensitivity_SE*1.96,normal_PA3_conf$Sensitivity_SE*1.96,normal_PA4_conf$Sensitivity_SE*1.96,normal_PA5_conf$Sensitivity_SE*1.96)
)

df1$Type = factor(c("Sift", "MCap", "CADD", "Revel", "noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)"), levels = c("Sift", "MCap", "CADD", "Revel", "noQC", "QC1" ,"QC2", "QC3", "QC3 (skipped QC2)"))

gg = ggplot(df1, aes(x=Sensitivity, y=Precision, colour = Type, shape = Type)) + geom_point(size=4) + 
  geom_errorbar(aes(ymin=Precision-Precision_CI, ymax=Precision+Precision_CI),colour="black",linetype=1,size=.2,width=.5) +
  geom_errorbarh(aes(xmin=Sensitivity-Sensitivity_CI, xmax=Sensitivity+Sensitivity_CI),colour="black",linetype=1,size=.2,height=0) +
  scale_shape_manual(values=c(0, 1, 2, 5, 15, 16, 17, 18, 19)) +
  xlim(0,1) + ylim(0,1)  + theme(plot.subtitle = element_text(vjust = 1),
    plot.caption = element_text(vjust = 1), 
    axis.title = element_text(size = 16), 
    plot.title = element_text(size = 14)) +labs(title = "Precision-Sensitivity space plot of Paralogue Annotation\ncompared with other tools", 
    x = "Sensitivity") + theme(axis.text = element_text(size = 13)) + theme(plot.title = element_text(hjust = 0.5)) 
ggsave("/media/nick/821ED5711ED55F2B/Users/Nick/Documents/PhD/Paralogues/ParalogueAnnotation_personal/images/PRC_of_TP53-TA_with_other_tools_plot.png")


p.normal_PA = Paralogous_var_align("data/UniFun/4704955/UniFun_deleterious.out_paraloc_paralogs2.noQC", "data/UniFun/4704955/UniFun_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA = Paralogous_var_align("data/UniFun/4704955/UniFun_benign.out_paraloc_paralogs2.noQC", "data/UniFun/4704955/UniFun_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
p.normal_PA2 = Paralogous_var_align("data/UniFun/4704955/UniFun_deleterious.out_paraloc_paralogs2.para_con", "data/UniFun/4704955/UniFun_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA2 = Paralogous_var_align("data/UniFun/4704955/UniFun_benign.out_paraloc_paralogs2.para_con", "data/UniFun/4704955/UniFun_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
p.normal_PA3 = Paralogous_var_align("data/UniFun/4704955/UniFun_deleterious.out_paraloc_paralogs2.all_con", "data/UniFun/4704955/UniFun_deleterious.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
b.normal_PA3 = Paralogous_var_align("data/UniFun/4704955/UniFun_benign.out_paraloc_paralogs2.all_con", "data/UniFun/4704955/UniFun_benign.out_paraloc_tableized", "data/clinvar/clinvar_20181028_GRCh37_onlyPathogenic_and_Likely_pathogenic.out_paraloc_tableized")
p.normal_PA4$Total_paralog_annotations = p.normal_PA3$Total_paralog_annotations[p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==p.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
b.normal_PA4$Total_paralog_annotations = b.normal_PA3$Total_paralog_annotations[b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.x==b.normal_PA3$Total_paralog_annotations$ALT_Amino_acids.y,]
normal_PA4 = Alt_con_PA(p.normal_PA4$Total_paralog_annotations,b.normal_PA4$Total_paralog_annotations)

other_tools = read.csv(file="data/UniFun/4704955/UniFun_deleterious.vcf.hg19_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
p.num_of_revel = length(which(other_tools$REVEL_score > 0.5))
p.num_of_sift = length(which(other_tools$SIFT_pred == "D"))
p.num_of_mcap = length(which(other_tools$M.CAP_pred == "D"))
p.num_of_cadd = length(which(other_tools$CADD_phred > 15))

b.other_tools = read.csv(file="data/UniFun/4704955/UniFun_benign.vcf.hg19_multianno.vcf_tableized_org", sep = "\t", header=TRUE, stringsAsFactors=FALSE) 
b.num_of_revel = length(which(b.other_tools$REVEL_score > 0.5))
b.num_of_sift = length(which(b.other_tools$SIFT_pred == "D"))
b.num_of_mcap = length(which(b.other_tools$M.CAP_pred == "D"))
b.num_of_cadd = length(which(b.other_tools$CADD_phred > 15))


# sift_conf = raw_conf_matrix(p.num_of_sift,other_tools,b.num_of_sift,b.other_tools)
# mcap_conf = raw_conf_matrix(p.num_of_mcap,other_tools,b.num_of_mcap,b.other_tools)
# revel_conf = raw_conf_matrix(p.num_of_revel,other_tools,b.num_of_revel,b.other_tools)
# cadd_conf = raw_conf_matrix(p.num_of_cadd,other_tools,b.num_of_cadd,b.other_tools)
sift1_conf = conf_matrix(p.num_of_sift,other_tools,b.num_of_sift,b.other_tools)
mcap1_conf = conf_matrix(p.num_of_mcap,other_tools,b.num_of_mcap,b.other_tools)
revel1_conf = conf_matrix(p.num_of_revel,other_tools,b.num_of_revel,b.other_tools)
cadd1_conf = conf_matrix(p.num_of_cadd,other_tools,b.num_of_cadd,b.other_tools)

normal_PA_conf = conf_matrix(p.normal_PA$true_num_of_paralog_anno, other_tools, b.normal_PA$true_num_of_paralog_anno, b.other_tools)
normal_PA2_conf = conf_matrix(p.normal_PA2$true_num_of_paralog_anno, other_tools, b.normal_PA2$true_num_of_paralog_anno, b.other_tools)
normal_PA3_conf = conf_matrix(p.normal_PA3$true_num_of_paralog_anno, other_tools, b.normal_PA3$true_num_of_paralog_anno, b.other_tools)
normal_PA4_conf = conf_matrix(normal_PA4$p.normal_PA4$true_num_of_paralog_anno, other_tools, normal_PA4$b.normal_PA4$true_num_of_paralog_anno, b.other_tools)

df2 = data.frame(
  Type = c("Sift","Mcap","Cadd","Revel","noQC","QC1","QC2","QC3"),
  Precision = c(sift1_conf$PPV,mcap1_conf$PPV,cadd1_conf$PPV,revel1_conf$PPV,normal_PA_conf$PPV,normal_PA2_conf$PPV,normal_PA3_conf$PPV,normal_PA4_conf$PPV),
  Precision_CI = c(sift1_conf$PPV_SE*.196,mcap1_conf$PPV_SE*1.96,cadd1_conf$PPV_SE*1.96,revel1_conf$PPV_SE*1.96,normal_PA_conf$PPV_SE*1.96,normal_PA2_conf$PPV_SE*1.96,normal_PA3_conf$PPV_SE*1.96,normal_PA4_conf$PPV_SE*1.96),
  Sensitivity = c(sift1_conf$Sensitivity,mcap1_conf$Sensitivity,cadd1_conf$Sensitivity,revel1_conf$Sensitivity,normal_PA_conf$Sensitivity,normal_PA2_conf$Sensitivity,normal_PA3_conf$Sensitivity,normal_PA4_conf$Sensitivity),
  Sensitivity_CI = c(sift1_conf$Sensitivity_SE*1.96,mcap1_conf$Sensitivity_SE*1.96,cadd1_conf$Sensitivity_SE*1.96,revel1_conf$Sensitivity_SE*1.96,normal_PA_conf$Sensitivity_SE*1.96,normal_PA2_conf$Sensitivity_SE*1.96,normal_PA3_conf$Sensitivity_SE*1.96,normal_PA4_conf$Sensitivity_SE*1.96)
)

```

# References










